---
title: 'Rasterize EEZs, MEOWs, and Bathymetry to Mollweide'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)
library(fasterize)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))
# source(here('spatial_fxns.R'))

```

# Summary

Set up analysis rasters for EEZs, MEOWs, and bathymetry at Gall-Peters 10 km^2 resolution.

* EEZs will include all EEZs as well as FAO regions and CCAMLR regions, based on the global OHI EEZ map.
* MEOWs will include all MEOW provinces, though only some are appropriate for IUCN regional assessments (see `2a_set_up_regional_codes.Rmd`).
* Bathymetry will include regions for 200m+, 20-200m, and <20m.  
* All will be clipped to the same ocean presence using the area raster.

# Data Sources

* Marine Ecoregions of the World: Spalding, M. D., Fox, H. E., Allen, G. R., Davidson, N., Ferdaña, Z. A., Finlayson, M. A. X., … others. (2007). Marine ecoregions of the world: A bioregionalization of coastal and shelf areas. BioScience, 57(7), 573–583.
* Bathymetry (GEBCO): Sandwell, D. T., Gille, S. T., & Smith, W. H. F. (2002, June). Bathymetry from Space:Oceanography, Geophysics, and Climate. Retrieved from https://www.gebco.net/

# Methods

## EEZ raster

Using the OHI global EEZ shapefile, we will include all EEZs by OHI region code, as well as FAO regions and Antarctica, using unique identifiers.  This will allow for separation by high seas vs. EEZ, or different takes on IUCN regional assessments using FAO or EEZ identifiers.

EEZ is rasterized for both land and sea, and then clipped to cells with non-NA values in the ocean_area raster.  This ensures presence of more coastal cells with fractional ocean area coverage.

``` {r set up EEZ raster}

eez_rast_file <- here('_spatial', 'eez_mol.tif')

area_rast <- raster(here('_spatial', 'ocean_area_mol.tif'))
values(area_rast)[values(area_rast) <= .005] <- NA

if(!file.exists(eez_rast_file)) {
  eez_shp_file  <- file.path(dir_M, 'git-annex/globalprep/spatial/v2017/regions_2017_update.shp')
  eez_sf <- read_sf(eez_shp_file)
  
  eez_df <- eez_sf %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  eez_ids <- eez_df %>%
    select(rgn_id, rgn_name, rgn_ant_id) %>%
    distinct()
  ### rgn_ant_id is same as rgn_id except for Antarctica.  We don't really
  ### need to chunk out Antarctica at this point so just use rgn_id.
  ### Filter out all land polygons and project to Gall-Peters
  ### EDIT: don't filter out land; leave it in, then later mask using the 
  ### area raster.  This should ensure more partial-ocean coastal cells are included.
  
  eez_sf_reclass <- eez_sf %>%
    clip_to_globe() %>%
    ### in common_fxn.R: ensures no coordinates outside +-180, +-90
    select(rgn_id, geometry) %>%
    st_transform(crs(area_rast))
  
  # plot(eez_sf_reclass)
  
  rast_base <- area_rast
  eez_rast <- fasterize::fasterize(eez_sf_reclass, rast_base, 
                                   field = 'rgn_id', fun = 'min')
  
  eez_rast_ocean <- eez_rast %>%
    mask(area_rast)
  
  writeRaster(eez_rast_ocean, eez_rast_file, overwrite = TRUE)
  
} else {
  # git_prov(eez_shp_file, 'input')
  # git_prov(eez_rast_file, 'output')
  eez_rast_ocean <- raster::raster(eez_rast_file)
}

plot(eez_rast_ocean, main = 'EEZ raster 10 km^2 Mollweide projection')

```

## MEOW raster

Using the MEOW shapefile, we will include all the shapes at the province level and then mask out land using the area raster (for consistency).

``` {r set up MEOW raster}

meow_rast_file <- here('_spatial', 'meow_mol.tif')
meow_shp_file  <- file.path(dir_bd_anx, 'meow/meow_ecos.shp')

if(!file.exists(meow_rast_file)) {
  meow_sf <- read_sf(meow_shp_file) %>%
    clip_to_globe()
    ### in common_fxn.R: ensures no coordinates outside +-180, +-90
  
  
  meow_sf_reclass <- meow_sf %>%
    select(ECO_CODE_X, geometry) %>%
    st_transform(crs(area_rast))
  
  # plot(meow_sf_reclass)
  
  rast_base <- area_rast
  
  meow_rast <- fasterize::fasterize(meow_sf_reclass, rast_base, 
                                   field = 'ECO_CODE_X', fun = 'min')
  
  meow_rast <- meow_rast %>%
    mask(area_rast)
    
  ### check whether there are shoreline gaps between MEOW maps and EEZ maps
  # x <- eez_rast
  # x[!is.na(values(meow_rast))] <- NA
  # plot(x)
  ### MEOWs seem to go inland to prevent shoreline gaps! yay
  
  writeRaster(meow_rast, meow_rast_file, overwrite = TRUE)
} else {
  # git_prov(meow_shp_file, 'input')
  # git_prov(meow_rast_file, 'output')
  meow_rast <- raster::raster(meow_rast_file)
}

plot(meow_rast, main = 'MEOW raster 10 km^2 Mollweide')

```


## Bathymetry raster

Using [General Bathymetric Chart of the Oceans (GEBCO) bathymetry raster](https://www.bodc.ac.uk/data/hosted_data_systems/gebco_gridded_bathymetry_data/), identify min depth of ocean cells.  For neritic and coastal cells, leaving as min depth allows for changing depth thresholds if desired.  We will not clip based on anything below 200 m, so can classify these more coarsely.

NOTE: 1 arcmin and 30 arcsec are available; 30 arcsec is more recent but one arcmin should be sufficient for our purposes.  Resample to 6 arc-minute (0.10°) using min depth in each cell.

Additionally, to ensure some minimum coastal cell presence, we will include a buffer based on the 3nm coastal zone from OHI.

Bathymetry is masked to the ocean area raster.  

### Create and view minimum depth raster

First, set up the coastal zone raster.  The original raster is 1/NA for whether a cell falls into the 3 nm coastal buffer, at 934 m resolution Mollweide.  Here, all NA cells are set to zero, non-NA remain 1, so aggregate will result in a proportional coastal zone coverage (so later we can select coastal cells based on some ocean presence threshold if desired).
```{r}
offshore_rast_10km_file <- here('_spatial', 'coastal_mol.tif')

if(!file.exists(offshore_rast_10km_file)) {
  offshore_rast_file <- file.path(dir_bd_anx, 'spatial', 'three_nm_offshore_mol.tif')
  
  x <- raster(offshore_rast_file)
  # values(x)[is.na(values(x))] <- 0
  y <- raster::aggregate(x, fact = 11, fun = sum, na.rm = TRUE, progress = 'text')
  z <- y / 121 ### divide by number of subcells in each supercell
  writeRaster(z, offshore_rast_10km_file, overwrite = TRUE)
}
```

Now we can use the coastal zone buffer to bolster the shallow water species presence to ensure some minimal inclusion even if the bathymetry drops off immediately.  The coastal zone buffer will also be useful when clipping land-based stressor layers.

``` {r set up bathymetry raster}

bathy_rast_file <- here('_spatial', 'bathy_mol.tif')

gebco_rast_file <- file.path(dir_bd_anx, 'spatial/gebco', 'GRIDONE_1D.nc')
gebco_rast_raw  <- raster(gebco_rast_file)

coastal_rast <- raster(offshore_rast_10km_file)

if(!file.exists(bathy_rast_file)) {
  ### The GEBCO raster is elevation, so depth is neg... for min depth, 
  ### aggregate by max (and vice versa)
  bathy_rast_min <- gebco_rast_raw %>%
    raster::aggregate(fact = 6, fun = max, progress = 'text')
  
  bathy_rast_min <- -1 * bathy_rast_min ### depth as positive
    
  rast_base <- coastal_rast
  
  bathy_rast_raw <- bathy_rast_min %>%
    projectRaster(rast_base, 
                  progress = 'text',
                  method = 'ngb')
  
  ### reclassify:
  ### compress cells above zero to sea level;
  ### compress cells below 200 m to rough depth clips
  bathy_rast <- bathy_rast_raw
  values(bathy_rast)[values(bathy_rast_raw) < 0] <- 0
  values(bathy_rast)[values(bathy_rast_raw) > 200] <- 1000
  values(bathy_rast)[values(bathy_rast_raw) > 1000] <- 2000
  values(bathy_rast)[values(bathy_rast_raw) > 2000] <- 3000
  values(bathy_rast)[values(bathy_rast_raw) > 3000] <- 4000
  values(bathy_rast)[values(bathy_rast_raw) > 4000] <- 5000
  values(bathy_rast)[values(bathy_rast_raw) > 5000] <- 6000
  values(bathy_rast)[values(bathy_rast_raw) > 6000] <- 7000
  values(bathy_rast)[values(bathy_rast_raw) > 7000] <- 8000
  
  bathy_rast <- bathy_rast %>%
    mask(area_rast)
    
  writeRaster(bathy_rast, bathy_rast_file, overwrite = TRUE)

  ### set up all masks with a base filled with NAs
  rast_neritic <- rast_shallow <- raster(bathy_rast)
  values(rast_neritic)[values(bathy_rast) <= 200] <- 1
  values(rast_neritic)[!is.na(values(coastal_rast))] <- 1
  values(rast_shallow)[values(bathy_rast) <=  60] <- 1
  values(rast_shallow)[!is.na(values(coastal_rast))] <- 1
  
  writeRaster(rast_neritic, here('_spatial', 'bathy_mol_neritic.tif'),
              overwrite = TRUE)
  writeRaster(rast_shallow, here('_spatial', 'bathy_mol_shallow.tif'),
              overwrite = TRUE)
  
} else {
  # git_prov(bathy_rast_file, 'output')
  bathy_rast <- raster::raster(bathy_rast_file)
}

plot(bathy_rast, main = 'Bathymetry (minimum depth) raster 10 km^2 Mollweide')

```

