---
title: 'Determine species ranges'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')  ###
library(sf)

source(here('common_fxns.R'))

```

# Summary

For the species rasterized in setup script 5b, determine their marine range using cell ocean areas from the ocean raster:
$$range = \sum_{cell = 1}^C A_{cell}$$

Save these ranges as an output to be used in the main scripts.  This will also include range per EEZ.

For those species listed with zero cells, check their polygon range to verify.  If polygon range is not zero, use this list to manually inspect polygons to make sure the rasterization process isn't dropping ranges.

# Data sources

* IUCN species shapefiles:  IUCN. (2020). The IUCN Red List of Threatened Species. Version 2020-1. Retrieved April 2020, from http://www.iucnredlist.org
* BirdLife International shapefiles: BirdLife International and Handbook of the Birds of the World. (2018). Bird species distribution maps of the world. Version 7.0. Available at http://datazone.birdlife.org/species/requestdis
* Bathymetry (GEBCO): Sandwell, D. T., Gille, S. T., & Smith, W. H. F. (2002, June). Bathymetry from Space:Oceanography, Geophysics, and Climate. Retrieved from https://www.gebco.net/


# Methods

## Read in each species rangemap (csv)

Bind the resulting dataframe with a dataframe of cell ocean areas.  Sum the total marine range.  Store total raster ranges for all spp in a dataframe.  Save this in the main intermediate directory for use in main scripts.

```{r}
range_from_rasts_file <- here('int/range_from_rasts.csv')
# unlink(range_from_rasts_file)
if(!file.exists(range_from_rasts_file)) {
  dir_mol_map <- file.path(dir_bd_anx, 'spp_rasts_mol_2020')
  
  ocean_a_rast <- raster(here('_spatial/ocean_area_mol.tif'))
  eez_rast     <- raster(here('_spatial/eez_mol.tif'))
  cell_km2     <- res(ocean_a_rast)[1]^2 / 1e6
  ocean_a_df   <- data.frame(cell_id = 1:ncell(ocean_a_rast),
                             eez     = values(eez_rast),
                             a_km2   = values(ocean_a_rast) * cell_km2) %>%
    filter(!is.na(a_km2))
    
  map_files <- list.files(dir_mol_map, full.names = TRUE)
  # f <- file.path(dir_mol_map, 'iucn_sid_10030.csv')
  # file.exists(f)

  range_l <- parallel::mclapply(map_files,
               FUN = function(f) {
                 ### f <- map_files[1]
                 id <- str_extract(basename(f), '[0-9]+') %>% as.integer()
                 x <- read_csv(f, col_types = 'ii') %>%
                   filter(presence > 0 & !is.na(presence) & presence != 5) %>%
                   dt_join(ocean_a_df, by = 'cell_id', type = 'left') 
                 
                 x_all <- x %>%
                   summarize(range_km2 = sum(a_km2, na.rm = TRUE)) %>%
                   mutate(eez = 0)
                 x_eez <- x %>%
                   group_by(eez) %>%
                   summarize(range_km2 = sum(a_km2, na.rm = TRUE)) %>%
                   ungroup()
                 
                 y <- bind_rows(x_all, x_eez) %>%
                   mutate(iucn_sid = id)
                 return(y)
               }, mc.cores = 40)
                              
  range_df <- bind_rows(range_l)
  
  write_csv(range_df, range_from_rasts_file)
}
```

## Check the ranges from the polygons

Read in the shapefile; calculate area of each of the polygons (`st_area()`) excluding extinct polygons (`filter(presence != 5)`).  Sum them for spp with multiple polygons.  NOTE: This doesn't account for overlapping polygons, but this is only intended as a quick sanity check - the polygon area is not used for any calculations.

Note also: this has many more species, but these are coming from non-marine birds, terrestrial mammals, terrestrial reptiles.

```{r}
range_from_polys_file <- here('_setup/int/range_from_polys.csv')

if(!file.exists(range_from_polys_file)) {
  
  dir_bli <- file.path(dir_M, 'git-annex/globalprep/_raw_data/birdlife_intl/d2019')
  dir_shp <- file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp/d2020-1')

  poly_files <- list.files(c(dir_bli, dir_shp), 
                           pattern = '.shp$', full.names = TRUE)
  
  range_df <- data.frame()
  
  for(poly_f in poly_files) {
    ### poly_f <- poly_files[1]
    message('Reading ', basename(poly_f))
    poly <- sf::read_sf(poly_f) 
    
    if(!'subpop' %in% names(poly)) {
      poly <- poly %>% mutate(subpop = NA)
    }
    if(!'iucn_sid' %in% names(poly)) {
      poly <- poly %>% rename(iucn_sid = id_no)
    }
    
    ids <- poly$iucn_sid %>% unique()

    poly_l <- parallel::mclapply(ids, 
                 FUN = function(id) { ### id <- ids[1]
                   message('Processing ', basename(poly_f), ' id: ', id)
                   x <- poly %>%
                     filter(iucn_sid == id) %>%
                     filter(presence != 5) %>%
                     mutate(poly_km2 = st_area(geometry) / 1e6)
                   y <- x %>% as.data.frame() %>%
                     group_by(iucn_sid, subpop) %>%
                     summarize(poly_km2 = sum(poly_km2, na.rm = TRUE)) %>%
                     mutate(poly_km2 = as.numeric(poly_km2))
                   
                   return(y)
                 }, mc.cores = 16)
    
    poly_df <- bind_rows(poly_l)
    range_df <- bind_rows(range_df, poly_df)
  }
  write_csv(range_df, range_from_polys_file)
}
```

## Combine 

```{r}
r_rast <- read_csv(range_from_rasts_file) %>%
  filter(eez == 0) %>%
  select(-eez)
r_poly <- read_csv(range_from_polys_file)
map_info <- read_csv(here('_data/spp_marine_maps_2020-1.csv'))
spp_risk <- read_csv(here('_data/iucn_risk_current_2020-1.csv')) %>%
  select(iucn_sid, cat, cat_score) %>%
  distinct()

x <- r_rast %>%
  left_join(map_info, by = 'iucn_sid') %>%
  left_join(r_poly, by = c('shp_iucn_sid' = 'iucn_sid', 'subpop')) %>%
  mutate(gp = tolower(basename(dbf_file)) %>% str_replace('.dbf', '')) %>%
  select(iucn_sid, subpop, sciname, gp, range_km2, poly_km2) %>%
  left_join(spp_risk, by = 'iucn_sid') %>%
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  distinct()

small_range_df <- x %>%
  filter(range_km2 < 10) %>%
  mutate(poly_km2 = as.integer(poly_km2)) %>%
  select(-subpop) %>%
  arrange(desc(poly_km2))

legit_small <- small_range_df %>%
  filter(poly_km2 < 100)

problem_small <- small_range_df %>%
  filter(poly_km2 >= 100)
```

### Check small-ranged species

`r knitr::kable(small_range_df)`

#### Checked (prev version):

Alosa caspia, Clupeonella caspia, Alosa braschnikowi, Alosa curensis, Alosa saposchnikowii, Clupeonella engrauliformis, Clupeonella grimmi, are Caspian sea fish.  Zero marine range is OK.  Excluded from assessment: 

135548, 135707, 98468449, 98468550, 98470302, 98471289, 98471433

#### Checked (current version):

For most of these, the polygon does not intersect what I have as coastal or neritic waters.  Could the polygons be displaced/shifted?  The Revillagigedo spp are near but not overlapping the islands identified in the ocean shapefile.  The St Paul's Rocks spp are dropped because neither the coastal buffer nor the bathymetry raster show shallow/coastal waters there (it's an archipelago about 500 feet across in the middle of the Atlantic).

* 22698097 Pterodroma caribbaea - Jamaican petrel polygon is entirely (far) inland.
* 178938	Enneanectes smithi - "Enneanectes smithi is known only from St. Paul's Rocks, Brazil in the central Atlantic. It has an area of occupancy of less than 5 km²." __VU__
* 183625	Halichoeres insularis	- "This species is endemic to the Eastern Pacific, and is only found in the found the Revillagigedo Islands. However, it may also be present at Alijos Rocks and Guadalupe Island, off the coast of central Baja." __VU__
* 184001	Labrisomus socorroensis	- "This species is endemic to the Eastern Pacific, and is only known from Revillagigedo Islands." __VU__


<!-- ### NA polygons? -->

<!-- Do any species show `NA` for polygon range... if so, shouldn't have a raster? -->

<!-- `r knitr::kable(small_range_df %>% filter(is.na(poly_km2)))` -->