---
title: 'Map stressor trends'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
# library(rgeos)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')  ###
  ### includes library(tidyverse); library(stringr); dir_M points to ohi directory

source(here('common_fxns.R'))
source(here('_setup/rasterize_fxns.R'))

```

# Summary

Using the stressor maps, aggregated to 10 km Mollweide, determine cell by cell linear trend.  This is recreating a script that was apparently lost in a big snafu.

# Data sources

Halpern et al. 2019

# Methods

## Make list of all stressor layers

Get all the stressor layers that have been aggregated to 10 km Mollweide.

```{r}
str_files_all <- list.files(file.path(dir_bd_anx, 'layers/stressors_10km'),
                        pattern = '.tif', full.names = TRUE)

str_df_all <- data.frame(file = str_files_all) %>%
  mutate(stressor = str_replace(basename(file), '_[0-9]{4}.+', ''),
         year = as.integer(str_extract(basename(file), '[0-9]{4}')))

str_years <- str_df_all %>%
  group_by(stressor) %>%
  summarize(year_min = min(year),
            year_max = max(year),
            range_03_13 = year_min <= 2003 & year_max >= 2013) %>%
  filter(range_03_13)

str_df <- str_df_all %>%
  filter(stressor %in% str_years$stressor) %>%
  filter(year >= 2003 & year <= 2013)
```

## Loop over all stressors, process linear model

For each stressor, across the limited time series, calculate per-cell trend in stressor value, including p value.

* load all years as a stack
* convert to a dataframe, dropping non-ocean cells
* for NA values, convert to zero - if NA means no presence of the stressor, then that is the same as a zero value?
* use mclapply to run a cell-wise linear model of stressor value by year
* get p value and the trend


```{r}

stressors <- str_df$stressor %>% unique() %>% sort()
ocean_rast <- raster(here('_spatial/ocean_area_mol.tif'))
ocean_a_df <- data.frame(cell_id = 1:ncell(ocean_rast),
                         ocean_prop = values(ocean_rast)) %>%
  filter(!is.na(ocean_prop))

outfile_dir <- file.path(dir_bd_anx, 'layers/stressor_trends')
# unlink(file.path(outfile_dir, '*.*'))
# unlink(file.path(outfile_dir, 'sst*.*'))

for(str in stressors) {
  # str <- stressors[14]
  message('Processing ', str)
  
  ### get the raw raster file for this stressor
  str_year_df <- str_df %>%
    filter(stressor == str)
  f <- str_year_df$file
  
  ### set up filename for the output
  outfile_stem <- file.path(outfile_dir, 'stressor_trend_%s_%s.tif')
  outfiles <- sprintf(outfile_stem, str, c('value', 'p_val'))
  
  if(any(!file.exists(outfiles))) {
    ### gather all rasters for this stressor into a stack
    str_stack <- raster::stack(f)
    
    str_rast_df_raw <- data.frame(values(str_stack)) %>%
      setNames(str_extract(names(.), '[0-9]{4}')) %>%
      mutate(cell_id = 1:ncell(str_stack[[1]])) %>%
      gather(year, val, -cell_id) 
    
    str_rast_df <- str_rast_df_raw %>%
      filter(cell_id %in% ocean_a_df$cell_id) %>%
      mutate(year = as.integer(year),
             val = ifelse(is.na(val), 0, val))
    
    cell_ids <- ocean_a_df$cell_id
    chunk_size <- 1000
    n_gps <- ceiling(length(cell_ids) / chunk_size)
    cell_gps <- rep(1:n_gps, each = chunk_size, length.out = length(cell_ids))
    
    trend_list <- parallel::mclapply(1:n_gps, mc.cores = 40,
      FUN = function(gp) { # gp <- 300 ### works well for clusters of 1000 cells
        if(gp %% 100 == 0) message('processing gp', gp)
        id_vec <- cell_ids[cell_gps == gp]
        
        cell_df <- str_rast_df %>% filter(cell_id %in% id_vec)
        
        cell_trend_df <- cell_df %>% 
          group_by(cell_id) %>%
          do(mdl = lm(val ~ year, data = .)) %>%
          broom::tidy(mdl) %>%
          ungroup() %>%
          filter(term == 'year') %>%
          select(cell_id, trend = estimate, p_val = p.value)
        return(cell_trend_df)
      })
 
    trend_df <- bind_rows(trend_list)
    val_rast <- map_to_rast(trend_df, cell_val = 'trend')
    pval_rast <- map_to_rast(trend_df, cell_val = 'p_val')

    writeRaster(val_rast, outfiles[1], overwrite = TRUE)
    writeRaster(pval_rast, outfiles[2], overwrite = TRUE)
    
  } else {
    message('Stressor trend files exist for stressor: ', str, '... skipping!')
  }
}
```

## Map results

```{r}
for(str in stressors) {
  ### str <- stressors[2]
  message('plotting trend layers for ', str)
  fstem <- file.path(dir_bd_anx, 'layers/stressor_trends/stressor_trend_%s_%s.tif')
  
  pval <- raster(sprintf(fstem, str, 'p_val'))
  val  <- raster(sprintf(fstem, str, 'value'))
  sig <- calc(pval, fun = function(x) x < .05)
  values(sig)[values(sig) < .1] <- NA
  
  high <- calc(val, fun = function(x) abs(x) > .001) %>%
    mask(sig)
  high1 <- calc(val, fun = function(x) abs(x) > .01) %>%
    mask(sig)

  plot(val, main = paste('Value: ', str))
  plot(sig, main = paste('Sig (p < .05): ', str))
  plot(high, main = paste('|Value| > .001 and p < .05: ', str),
       col = c('red', 'darkgreen'))
  plot(high1, main = paste('|Value| > .01 and p < .05: ', str),
       col = c('red', 'darkgreen'))
}
```

