---
title: "Combine stressor sensitivities"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  word_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    highlight: haddock
    includes:
      in_header: ~/github/src/templates/ohara_hdr.html
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here::here('common_fxns.R'))
library(flextable)
```


# Summary

Stressor sensitivity is a very simplified version of stressor weights (see draft1 folder for complicated version).  Here, we will simply note whether a given species is sensitive to a given stressor, using the IUCN impact weights (i.e. non-negligible impact score), ranked as 0-3 (no, low, medium, high impact) for later differentiation if necessary.  For species with NA impact scores, we can imply sensitivity based on other fields, e.g. severity.  This version of sensitivity is not related to the range of exposure, as the weight calculation was doing.

Sensitivity will denote whether a species range overlapping a stressor range constitutes an impact.

# Methods

Gather processed sensitivity dataframe.

```{r}

spp_incl <- get_incl_spp()

```

## Match threat codes to stressors

```{r}
threat_to_stressor <- read_csv(here('_raw/iucn_threat_to_stressor_lookup.csv')) %>%
  mutate(code_main = str_extract(code, '[0-9]+') %>% as.numeric(),
         code_short = str_extract(code, '[0-9]+\\.[0-9]+')) %>%
  mutate(stressor = str_split(stressor, ';')) %>%
  unnest(stressor) %>%
  separate(desc, into = c('desc_main', 'desc_sub', 'desc_subsub'), 
           sep = ': ', remove = FALSE)

thr_to_spp <- spp_incl %>%
  left_join(threat_to_stressor, by = c('code', 'stressor', 'category')) %>%
  mutate(desc_main = sprintf('%s. %s', code_main, desc_main),
         desc_sub = sprintf('%s. %s', code_short, desc_sub)) %>%
  select(code_main, desc_main, desc_sub, stressor, iucn_sid, category) %>%
  filter(!is.na(stressor)) %>%
  distinct()

threat_table <- thr_to_spp %>%
  group_by(code_main, desc_main, desc_sub, stressor)  %>%
  summarize(n_spp = n_distinct(iucn_sid)) %>%
  ungroup() %>%
  arrange(code_main, stressor) %>%
  select(-code_main) %>%
  setNames(c('Threat category', 'Threat subcategory', 'Stressor', 'n spp'))

flextable(threat_table) %>%
  theme_vanilla() %>%
  align(part = 'all', align = 'left') %>%
  align(part = 'all', j = 4, align = 'center') %>%
  # bg(bg = 'white', part = 'header') %>%
  # border(border.bottom = officer::fp_border(), part = 'header') %>%
  font(font = 'Arial', part = 'all') %>%
  fontsize(size = 8, part = 'all') %>%
  width(1:2, width = 2.5) %>%
  width(3, width = 1) %>%
  merge_v(j = 1:3)
  
```

