---
title: 'Calculate land- and ocean-based stressor sensitivities from IUCN impact scores'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here::here('common_fxns.R'))
```


# Summary

Stressor sensitivity is a very simplified version of stressor weights (see draft1 folder for complicated version).  Here, we will simply note whether a given species is sensitive to a given stressor, using the IUCN impact weights (i.e. non-negligible impact score), ranked as 0-3 (no, low, medium, high impact) for later differentiation if necessary.  For species with NA impact scores, we can imply sensitivity based on other fields, e.g. severity.  This version of sensitivity is not related to the range of exposure, as the weight calculation was doing.

Sensitivity will denote whether a species range overlapping a stressor range constitutes an impact.

# Methods

## Set up stressor sensitivity dataframe

Using all listed threats and the threat-to-stressor lookup, identify species with non-negligible impacts from impact score, and where impact is unknown, based on severity field.  Here we will consider all species, regardless of mapped status, comprehensively assessed status, or threatened status.

In cases where a species is impacted by multiple threats that map to a single stressor (e.g. different activities that map to "direct human" impact) we summarize to the _maximum_ score out of all listed.  For example, a species suffers low impact from residential development (score = 3) but medium impact from commercial (score = 5), the direct human impact layer will be weighted as a 5 across the board.

```{r set up spp with threats}
spp_marine <- read_csv(here('_data', sprintf('spp_marine_from_api_%s.csv', api_version)))

spp_maps <- read_csv(here('_data', sprintf('spp_marine_maps_%s.csv', api_version)))

spp_risk <- read_csv(here('_data', sprintf('iucn_risk_current_%s.csv', api_version))) %>%
  select(iucn_sid, cat_score)

spp_comp <- read_csv(here('_data', sprintf('iucn_comp_assessed_%s.csv', api_version)))

```

### Calc for land-based stressors

Nutrient, chemical, direct human, and light

``` {r summarize sensitivity to land stressors}
chi_lookup <- read_csv(here('_raw/iucn_threat_to_stressor_lookup.csv')) %>%
  mutate(stressor = str_split(stressor, ';')) %>%
  unnest(stressor) %>%
  filter(str_detect(stressor, '[a-z]')) %>%
  filter(category == 'land-based')

spp_threats <- read_csv(file.path(dir_bd_anx, 'iucn/threats',
                               sprintf('iucn_spp_threats_%s.csv', api_version))) %>%
  left_join(chi_lookup, by = 'code') %>%
  filter(!is.na(code))

### Join the marine species list to the threats.
### * recategorize sensitivities based on impact score
### * if no impact score, categorize sensitivities based on severity
### * rather than filtering, set NA and past scores to 0
###   * this allows non-threatened species to remain in the mix
threats_to_marine <- spp_marine %>%
  left_join(spp_risk, by = 'iucn_sid') %>%
  left_join(spp_threats, by = 'iucn_sid') %>%
  ### code sensitivity based on impact:
  mutate(sens = case_when(score_num %in% 0:2 ~ FALSE,   ### no/negligible impact
                          score_num > 2      ~ TRUE,
                          TRUE               ~ NA)) %>% ### set all else to NA
  ### for NA impact, code sensitivity based on severity:
  mutate(sev = tolower(severity),
         sens = case_when(is.na(sens) & str_detect(sev, 'negligible|no decl') ~ FALSE,
                          is.na(sens) & str_detect(sev, 'very rapid decl')    ~ TRUE,
                          is.na(sens) & str_detect(sev, 'rapid decl')         ~ TRUE,
                          is.na(sens) & str_detect(sev, 'causing|slow, significant') ~ TRUE,
                          TRUE ~ sens)) %>%
  ### set past impacts and unresolved sensitivies to 0:
  mutate(sens = ifelse(timing == 'Past, Unlikely to Return', FALSE, sens),
         sens = ifelse(is.na(sens), FALSE, sens)) %>%
  filter(!is.na(code)) %>% ### drop code mismatches
  filter(!is.na(stressor)) %>% ### drop codes that don't match stressors
  select(iucn_sid, cat_score,
         sens, stressor, code, score_num) %>%
  distinct()

### sensitivity categories:
# [1] "Unknown"                          "Negligible declines"             
# [3] NA                                 "No decline"                      
# [5] "Rapid Declines"                   "Slow, Significant Declines"      
# [7] "Causing/Could cause fluctuations" "Very Rapid Declines"   

write_csv(threats_to_marine %>% select(-cat_score),
          here('int', sprintf('spp_sens_land_%s.csv', api_version)))
```

``` {r calc comparisons land}
n_comp <- spp_marine %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  .$iucn_sid %>% unique() %>% length()
n_threatened <- spp_risk %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  .$iucn_sid %>% unique() %>% length()

n_stressed <- threats_to_marine %>%
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  filter(sens) %>%
  .$iucn_sid %>% unique() %>% length()

```

Out of the list of `r n_comp` marine species with maps in comprehensively-assessed taxa, `r n_threatened` are not Least Concern, Extinct, nor Data Deficient.  

The resulting number of threatened spp (including Near Threatened) with sensitivity to at least one __land-based__ stressor is: `r n_stressed`.  Land-based stressors comprise:

* `direct_human`
* `nutrient`
* `organic`
* `light`

NOTE: These numbers may include a small handful of spp with effectively zero ocean range, which will not show up in maps, will not affect calculations, etc.  These can be found from the impacted area dataframe.
<!-- 6957 comp-assessed mapped species  -->
<!-- 1357 spp with category NC to CR  -->
<!--  631 with sensitivity to land-based stressors -->

### Calc for ocean-based stressors

Just shipping...

``` {r summarize sensitivity to ocean stressors}
chi_lookup <- read_csv(here('_raw/iucn_threat_to_stressor_lookup.csv')) %>%
  mutate(stressor = str_split(stressor, ';')) %>%
  unnest(stressor) %>%
  filter(str_detect(stressor, '[a-z]')) %>%
  filter(category == 'ocean')

spp_threats <- read_csv(file.path(dir_bd_anx, 'iucn/threats',
                               sprintf('iucn_spp_threats_%s.csv', api_version))) %>%
  left_join(chi_lookup, by = 'code') %>%
  filter(!is.na(code))

### Join the marine species list to the threats.
### * recategorize sensitivities based on impact score
### * if no impact score, categorize sensitivities based on severity
### * rather than filtering, set NA and past scores to 0
###   * this allows non-threatened species to remain in the mix
threats_to_marine <- spp_marine %>%
  left_join(spp_risk, by = 'iucn_sid') %>%
  left_join(spp_threats, by = 'iucn_sid') %>%
  ### code sensitivity based on impact:
  mutate(sens = case_when(score_num %in% 0:2 ~ FALSE,   ### no/negligible impact
                          score_num > 2      ~ TRUE,
                          TRUE               ~ NA)) %>% ### set all else to NA
  ### for NA impact, code sensitivity based on severity:
  mutate(sev = tolower(severity),
         sens = case_when(is.na(sens) & str_detect(sev, 'negligible|no decl') ~ FALSE,
                          is.na(sens) & str_detect(sev, 'very rapid decl')    ~ TRUE,
                          is.na(sens) & str_detect(sev, 'rapid decl')         ~ TRUE,
                          is.na(sens) & str_detect(sev, 'causing|slow, significant') ~ TRUE,
                          TRUE ~ sens)) %>%
  ### set past impacts and unresolved sensitivies to 0:
  mutate(sens = ifelse(timing == 'Past, Unlikely to Return', FALSE, sens),
         sens = ifelse(is.na(sens), FALSE, sens)) %>%
  filter(!is.na(code)) %>% ### drop code mismatches
  filter(!is.na(stressor)) %>% ### drop codes that don't match stressors
  select(iucn_sid, cat_score,
         sens, stressor, code) %>%
  distinct()

### sensitivity categories:
# [1] "Unknown"                          "Negligible declines"             
# [3] NA                                 "No decline"                      
# [5] "Rapid Declines"                   "Slow, Significant Declines"      
# [7] "Causing/Could cause fluctuations" "Very Rapid Declines"   

write_csv(threats_to_marine %>% select(-cat_score),
          here('int', sprintf('spp_sens_ocean_%s.csv', api_version)))
```

``` {r calc comparisons ocean}
n_comp <- spp_marine %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  .$iucn_sid %>% unique() %>% length()
n_threatened <- spp_risk %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  .$iucn_sid %>% unique() %>% length()

n_stressed <- threats_to_marine %>%
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  filter(sens) %>%
  .$iucn_sid %>% unique() %>% length()

```

Out of the list of `r n_comp` marine species with maps in comprehensively-assessed taxa, `r n_threatened` are not Least Concern, Extinct, nor Data Deficient.  

The resulting number of threatened spp (including Near Threatened) with sensitivity to at least one __ocean-based__ stressor is: `r n_stressed`.  Ocean-based stressors comprise:

* `shipping`

NOTE: These numbers may include a small handful of spp with effectively zero ocean range, which will not show up in maps, will not affect calculations, etc.  These can be found from the impacted area dataframe.
<!-- 6957 comp-assessed mapped species  -->
<!-- 1357 spp with category NC to CR  -->
<!--  438 with sensitivity to general stressors -->
