---
title: 'Calculate climate stressor sensitivities from IUCN impact scores'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here::here('common_fxns.R'))

library(tidytext) ### to help parse ngrams in the narratives

```


# Summary

Stressor sensitivity is a very simplified version of stressor weights (see draft1 folder for complicated version).  Here, we will simply note whether a given species is sensitive to a given stressor, using the IUCN impact weights (i.e. non-negligible impact score), ranked as 0-3 (no, low, medium, high impact) for later differentiation if necessary.  For species with NA impact scores, we can imply sensitivity based on other fields, e.g. severity.  This version of sensitivity is not related to the range of exposure, as the weight calculation was doing.

Sensitivity will denote whether a species range overlapping a stressor range constitutes an impact.

# Methods

## Set up stressor sensitivity dataframe

Using all listed threats and the threat-to-stressor lookup, and focusing on those species with valid maps (i.e. a csv file in `spp_rasts_mol_2019`), identify species with non-negligible impacts from impact score, and where impact is unknown, based on severity field.  Here we are looking at climate threats.  To differentiate between SST, SLR, OA, and possibly UV sensitivity (though UV not included in Halpern 2019), we will mine the species narratives for keywords, and use habitat information to locate the species in the water column.

In cases where a species is impacted by multiple threats that map to a single stressor (e.g. different mentions of SST in different threat categories) we summarize to the _maximum_ score out of all listed.  For example, a species suffers low impact from residential development (score = 3) but medium impact from commercial (score = 5), the direct human impact layer will be weighted as a 5 across the board.

In this step we will include all species including LC and DD.  We will however limit to comprehensively assessed taxa.

| Code  | Description                                   | Potential strs  |
|:-----:|:----------------------------------------------|:----------------|
|   11  | Climate change & severe weather               | NA              |
|  11.1 | Habitat shifting & alteration                 | sst, slr, oa    | 
|  11.2 | Droughts                                      | NA              | 
|  11.3 | Temperature extremes                          | sst, oa^*^      | 
|  11.4 | Storms & flooding                             | NA              | 
|  11.5 | Other impacts                                 | sst^*^, oa, slr | 

\* Note that ocean acidification effects are often described in narratives for species only classified with code 11.3, and sea surface temperature, ocean acidification, and sea level rise impacts are often described in narratives for species only classified with code 11.5.

```{r set up spp with maps and threats, eval = TRUE}
spp_marine <- read_csv(here('_data', sprintf('spp_marine_from_api_%s.csv', api_version)))

spp_maps <- read_csv(here('_data', sprintf('spp_marine_maps_%s.csv', api_version)))

spp_risk <- read_csv(here('_data', sprintf('iucn_risk_current_%s.csv', api_version))) %>%
  select(iucn_sid, cat_score)

spp_comp <- read_csv(here('_data', sprintf('iucn_comp_assessed_%s.csv', api_version)))

chi_lookup <- read_csv(here('_raw/iucn_threat_to_stressor_lookup.csv')) %>%
  mutate(stressor = str_split(stressor, ';')) %>%
  unnest(stressor) %>%
  filter(str_detect(stressor, '[a-z]')) %>%
  filter(category == 'climate')

spp_threats <- read_csv(file.path(dir_bd_anx, 'iucn/threats',
                               sprintf('iucn_spp_threats_%s.csv', api_version))) %>%
  left_join(chi_lookup, by = 'code') %>%
  filter(!is.na(code))

### Join the maps to the threats.
### * recategorize sensitivities based on impact score
### * if no impact score, categorize sensitivities based on severity
### * rather than filtering, set NA and past scores to 0
###   * this allows non-threatened species to remain in the mix
threats_to_marine <- spp_marine %>%
  left_join(spp_risk, by = 'iucn_sid') %>%
  left_join(spp_threats, by = 'iucn_sid') %>%
  ### code low-med-high sensitivity based on impact:
  mutate(sens = case_when(score_num %in% 0:2 ~ FALSE,     ### no/negligible impact
                          score_num > 2      ~ TRUE,      ### non-negligible impact
                          TRUE               ~ NA)) %>%   ### set all else to NA
  ### for NA impact, code sensitivity based on severity:
  mutate(sev = tolower(severity),
         sens = case_when(is.na(sens) & str_detect(sev, 'negligible|no decl') ~ FALSE,
                          is.na(sens) & str_detect(sev, 'very rapid decl')    ~ TRUE,
                          is.na(sens) & str_detect(sev, 'rapid decl')         ~ TRUE,
                          is.na(sens) & str_detect(sev, 'causing|slow, significant') ~ TRUE,
                          TRUE ~ sens)) %>%
  ### set past impacts and unresolved sensitivies to 0:
  mutate(sens = ifelse(timing == 'Past, Unlikely to Return', FALSE, sens),
         sens = ifelse(is.na(sens), FALSE, sens)) %>%
  filter(!is.na(code)) %>% ### drop code mismatches
  filter(!is.na(stressor)) %>% ### drop codes that don't match stressors
  select(iucn_sid, cat_score, sens, 
         code, stressor) %>%
  distinct()

### sensitivity categories:
# [1] "Unknown"                          "Negligible declines"             
# [3] NA                                 "No decline"                      
# [5] "Rapid Declines"                   "Slow, Significant Declines"      
# [7] "Causing/Could cause fluctuations" "Very Rapid Declines"   

write_csv(threats_to_marine, here('int/spp_threats_climate.csv'))

```

``` {r}
n_comp <- spp_marine %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  .$iucn_sid %>% unique() %>% length()
n_threatened <- spp_risk %>% 
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  .$iucn_sid %>% unique() %>% length()

spp_stressed <- threats_to_marine %>%
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  filter(sens) 

n_stressed <- spp_stressed %>%
  .$iucn_sid %>% unique() %>% length()
  
```

Out of the list of `r n_comp` marine species with maps, `r n_threatened` are neither Least Concern nor Data Deficient.  

The resulting number of threatened spp (including Near Threatened) with sensitivity to at least one __climate__ stressor is: `r n_stressed`.

NOTE: These numbers may include a small handful of spp with effectively zero ocean range, which will not show up in maps, will not affect calculations, etc.  These can be found from the impacted area dataframe.
<!-- 6957 comp-assessed mapped species -->
<!-- 1357 spp with category NC to CR -->
<!-- 555 with sensitivity to climate stressors -->

## Identify species habitat location (shallow vs. deep)

For fishing stressors, we differentiated between demersal and pelagic to predict likely impacts from different gear types.  Here for climate impacts we examine shallow vs. deep species to help parse potential sensitivity to sea surface temperature and UV exposure, and species who rely on intertidal and shore to parse sensitivity to sea level rise.  

* For SST: Shallow habitats include all neritic (9.X), epipelagic (10.1), marine intertidal (12.X), and marine coastal/supratidal (13.X).
* For SLR: Shoreline habitats include marine intertidal (12.X) and coastal/supratidal (13.X)
    * consider perhaps reef, seagrass, and estuary habitats (9.8, 9.9, 9.10 respectively). These are not currently included.

```{r}
spp_shallow_shore <- read_csv(file.path(dir_bd_anx, 'iucn',
                                sprintf('spp_habs_from_api_%s.csv', api_version)),
                     col_types = cols('iucn_sid' = 'i',
                                      'code' = 'c')) %>%
  select(iucn_sid, code, habitat, suitability) %>%
  distinct() %>%
  mutate(code_main = as.integer(str_replace_all(code, '\\..*', ''))) %>%
  filter(code_main %in% 9:13) %>% ### limit to marine spp
  mutate(shallow   = code_main %in% c(9, 12, 13) | code == '10.1',
         shoreline = code_main %in% c(12, 13)) %>%
  group_by(iucn_sid) %>%
  summarize(shallow = any(shallow),
            shoreline = any(shoreline)) %>%
  ungroup()

write_csv(spp_shallow_shore, here('int/spp_shallow_shore.csv'))

```

## Check for climate instances in narratives

### identify keywords for sea surface temperature

Examine trigrams from the narratives that include 'temp' or 'sst'.  Limit to the "threats" dimension since temperature may show up in "habitat" or other dimensions more innocuously.

```{r set up for ngram analysis}

spp_narr_file <- file.path(dir_bd_anx, 'iucn', 
                           sprintf('spp_narr_from_api_%s.csv', api_version))

spp_narrs <- read_csv(spp_narr_file,
                      col_types = cols(.default = 'c', iucn_sid = 'i')) %>%
  select(iucn_sid:usetrade) %>%
  gather(dimension, text, -iucn_sid) %>%
  ### remove HTML tags since they seem to affect the unnesting of sentences
  mutate(text = str_replace_all(text, '<.*?>', ' '))

### set up some stop words.  The stop_words df is a bit large and slow
### to search over... also doesn't drop numbers
cut_words <- data.frame(
    word = c('a', 'and', 'or', 'in', 'on', 'of', 'by',
             'to', 'but', 'the', 'from', 
             '[0-9,;\\.]+')
  ) %>%
  mutate(combo = paste0('^', word, ' | ', word, '$')) %>%
    ### e.g. not "and trawl" or "trawling and"
  summarize(to_cut = paste0(combo, collapse = '|'))

```

```{r examine temperature trigrams}

sst_search_string <- 'temperature|sst'

sst_text_df <- spp_narrs %>%
  filter(dimension == 'threats') %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = FALSE) %>%
  ### prefilter to dramatically speed up the trigrams...
  filter(str_detect(tolower(sentence), sst_search_string)) %>%
  # group_by(iucn_sid, dimension) %>%
  unnest_tokens(output = trigram, input = sentence, 
                token = 'ngrams', n = 3, drop = FALSE) %>%
  filter(str_detect(trigram, sst_search_string))


sst_trigram_df <- sst_text_df %>%
  select(-text, -iucn_sid) %>%
  filter(!str_detect(trigram, cut_words$to_cut)) %>%
  group_by(trigram) %>%
  summarize(instances = n()) %>%
  ungroup() %>%
  arrange(desc(instances)) %>%
  filter(instances > 1)

DT::datatable(sst_trigram_df)

```

Some frequent word combos:

* sea surface temperature
* increased (ocean/water)? temperature, increasing (water/sea)? temperature
* high (sea)? temperature
* temperature extreme, extreme temperature events, extreme elevated temperature
* temperatures willis (particularly for corals)
    * based on "Increased coral disease levels on the GBR were correlated with increased ocean temperatures (Willis et al. 2007) supporting the prediction that disease levels will be increasing with higher sea surface temperatures"
* caution to avoid sand or air temperature combinations!

### identify keywords for sea ice loss

A number of mammals and sea birds are sensitive to sea ice loss, dependent on the ice for reproduction and pupping, haul outs, and protection.  Here examine the full suite of spp to see if other taxa have sea ice dependency (and to avoid it if indirect).

```{r examine sea ice trigrams}

ice_search_string <- 'sea ice'

ice_text_df <- spp_narrs %>%
  filter(dimension == 'threats') %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = FALSE) %>%
  ### prefilter to dramatically speed up the trigrams...
  filter(str_detect(tolower(sentence), ice_search_string)) %>%
  # group_by(iucn_sid, dimension) %>%
  unnest_tokens(output = trigram, input = sentence, 
                token = 'ngrams', n = 3, drop = FALSE) %>%
  filter(str_detect(trigram, ice_search_string))


ice_trigram_df <- ice_text_df %>%
  filter(!str_detect(trigram, cut_words$to_cut)) %>%
  group_by(trigram) %>%
  summarize(instances = n()) %>%
  ungroup() %>%
  arrange(desc(instances)) %>%
  filter(instances > 1)

DT::datatable(ice_trigram_df)

```

Inspecting the results, `sea ice` on its own appears to capture legitimate threats.

### identify keywords for ENSO and global warming

Many shallow water spp e.g. blennies and wrasses indicate a threat from water temperatures/oceanographic changes due to ENSO and global warming.  Many (but not all) use similar phrasing: 

* "Given this species' restricted distribution and shallow water habitat, oceanographic environmental changes, such as those associated with future ENSO events, may have detrimental effects on the survival of this species."
* "Given this species' potentially restricted distribution and shallow water habitat, localized stochastic events, including future oceanographic environmental changes such as ENSO and global warming, may have detrimental effects on marine ecosystems and the survival of these shallow water regional endemic species."

```{r examine enso trigrams}

enso_search_string <- '(^|[^a-z])enso($|[^a-z])|global warming'
  ### the regex keeps from 'enso' cueing on 'Tittensor' e.g.

enso_text_df <- spp_narrs %>%
  filter(dimension == 'threats') %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = FALSE) %>%
  ### prefilter to dramatically speed up the trigrams...
  filter(str_detect(tolower(sentence), enso_search_string)) %>%
  # group_by(iucn_sid, dimension) %>%
  unnest_tokens(output = trigram, input = sentence, 
                token = 'ngrams', n = 3, drop = FALSE) %>%
  filter(str_detect(trigram, enso_search_string))


enso_trigram_df <- enso_text_df %>%
  filter(!str_detect(trigram, cut_words$to_cut)) %>%
  group_by(trigram) %>%
  summarize(instances = n()) %>%
  ungroup() %>%
  arrange(desc(instances)) %>%
  filter(instances > 1)

DT::datatable(enso_trigram_df)

```

Inspection of the text indicates that `enso` and `global warming` both capture legitimate instances of threats.

### identify keywords for ocean acidification

Examine trigrams for instances of "ocean acid".

```{r examine OA trigrams}

oa_search_string <- 'ocean acidification|ocean acidity'
### 1077 caught by 'acid' alone, 1055 caught by 'ocean acid',
### 1053 caught by 'ocean acidification'
### dropped instances reference domoic acid or lead acid batteries.

oa_text_df <- spp_narrs %>%
  filter(dimension == 'threats') %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = FALSE) %>%
  ### prefilter to dramatically speed up the trigrams...
  filter(str_detect(tolower(sentence), oa_search_string)) %>%
  # group_by(iucn_sid, dimension) %>%
  unnest_tokens(output = trigram, input = sentence, 
                token = 'ngrams', n = 3, drop = FALSE) %>%
  filter(str_detect(trigram, oa_search_string))

oa_trigram_df <- oa_text_df %>%
  select(-text, -iucn_sid) %>%
  filter(!str_detect(trigram, cut_words$to_cut)) %>%
  group_by(trigram) %>%
  summarize(instances = n()) %>%
  ungroup() %>%
  arrange(desc(instances)) %>%
  filter(instances > 1)

DT::datatable(oa_trigram_df)

```

It appears that "ocean acid" on its own is a decent search string to identify ocean acidification threats.

### identify keywords for sea level rise

Examine tetragrams for instances of "sea level rise" or similar.

```{r examine SLR tetragrams}

slr_search_string <- 'sea level'
### 1077 caught by 'acid' alone, 1055 caught by 'ocean acid',
### 1053 caught by 'ocean acidification'
### dropped instances reference domoic acid or lead acid batteries.

slr_text_df <- spp_narrs %>%
  filter(dimension == 'threats') %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = FALSE) %>%
  ### prefilter to dramatically speed up the trigrams...
  filter(str_detect(tolower(sentence), slr_search_string)) %>%
  # group_by(iucn_sid, dimension) %>%
  unnest_tokens(output = tetragram, input = sentence, 
                token = 'ngrams', n = 4, drop = FALSE) %>%
  filter(str_detect(tetragram, slr_search_string))
### n_ids(slr_text_df)
### no hits on 'slr'; 134 hits with just 'sea level'; 65 with 'sea level rise

slr_tetragram_df <- slr_text_df %>%
  select(-text, -iucn_sid) %>%
  filter(!str_detect(tetragram, cut_words$to_cut)) %>%
  group_by(tetragram) %>%
  summarize(instances = n()) %>%
  ungroup() %>%
  arrange(desc(instances)) %>%
  filter(instances > 1)

DT::datatable(slr_tetragram_df)

```

Some potential matches:

* sea level rise/rising
* changing/rising sea levels
* sea level change/changing

## Parse threats to count by climate stressor type

Using the above analyses, set up keywords to map species sensitivities to the various climate stressors.

```{r set up climate stressor keywords}
### set up search terms for sea level rise:
slr_keywords <- c('sea level ris', 'changing sea level', 'rising sea level', 'sea level chang') %>%
  paste(collapse = '|')

### set up search terms for ocean acidification:
oa_keywords <- 'ocean acidifi'

### set up search terms for sea surface temperature:
sst_keywords <- c('sea surface temper',
                  'increas[a-z]+ (ocean|water|sea)? ?temper',
                  'high sea temper',
                  'temper[a-z]+ extreme',
                  'extreme (elevated )?temper',
                  'climate change',
                  'el ni[nñ]o', ### the ñ is not working right...
                  '(^|[^a-z])enso($|[^a-z])',
                  'sea ice',
                  'global warming',
                  'temper[a-z]+ willis') %>%
  paste0(collapse = '|')

cc_keywords <- paste(slr_keywords, oa_keywords, sst_keywords, sep = '|')

```


``` {r search narratives for terms}

cc_narr_df <- spp_narrs %>%
  filter(dimension %in% c('threats', 'rationale')) %>%
  tidytext::unnest_tokens(input = text, output = sentence, 
                          token = 'sentences', drop = TRUE) %>%
  filter(str_detect(sentence, cc_keywords)) %>%
  mutate(oa_count  = str_count(sentence, oa_keywords),
         slr_count = str_count(sentence, slr_keywords),
         sst_count = str_count(sentence, sst_keywords))

cc_narr_tidy <- cc_narr_df %>%
  gather(stressor, count, ends_with('count')) %>%
  mutate(stressor = str_replace_all(stressor, '_count', '')) %>%
  group_by(iucn_sid, stressor) %>%
  summarize(count = sum(count))

# x <- cc_narr_tidy %>%
#   filter(stressor == 'oa' & count > 0)
### 1060 OA spp still here

write_csv(cc_narr_tidy, here('int/cc_narr_tidy.csv'))

```

## Combine stressors with species IDs and habitats

Note that here we filter out occurrences of:

* species whose narrative suggests sensitivity to sea surface temperature but whose habitat does not include surface waters (<200m)
* species whose narrative suggests sensitivity to sea level rise but whose habitat does not include shoreline habitats.

```{r}
cc_narr_tidy          <- read_csv(here('int/cc_narr_tidy.csv'))
spp_shallow_shore     <- read_csv(here('int/spp_shallow_shore.csv'))
spp_threats_from_iucn <- read_csv(here('int/spp_threats_climate.csv'))

### identify all the unique climate stressors... drop 'uv' for now
# stressors <- spp_threats_from_iucn$stressor %>% unique()
stressors <- c('sst', 'slr', 'oa')

### join stressors from narratives to impact scores; filter out
### stressors that don't match with habitats.
cc_sens_df <- cc_narr_tidy %>%
  left_join(spp_shallow_shore, by = 'iucn_sid') %>%
  left_join(spp_threats_from_iucn, by = c('iucn_sid', 'stressor')) %>%
  filter(!(stressor == 'sst' & shallow == FALSE)) %>%
  filter(!(stressor == 'slr' & shoreline == FALSE))
### 866 spp sensitive to SST without regard to habitat; 865 after filter
### 75 spp sensitive to SLR without regard to habitat; 74 after filter.
### only 10 spp sensitive to ocean acidification - that does not seem right!
### Check that in the next code chunk.

### Summarize sensitivity to each stressor as any non-zero counts
### within that stressor category.  Complete the list so all stressors
### show up for all species, leaving NAs for those added via complete().
cc_sens_summary <- cc_sens_df %>%
  group_by(iucn_sid, stressor, code) %>%
  filter(!is.na(sens)) %>%
  summarize(sens = sum(count * sens) > 0) %>%
  group_by(iucn_sid) %>%
  complete(stressor = stressors) %>%
  ungroup()

write_csv(cc_sens_summary, 
          here('int', sprintf('spp_sens_climate_%s.csv', api_version)))
```

<!-- ### Why so few instances of sensitivity to ocean acidification? -->

<!-- NOTE: THIS HAS BEEN RESOLVED, see end of commented out section. -->

<!-- There are many corals included in the dataset, which do not seem to be caught by this process as being sensitive to ocean acidification.  The threat narratives claim ocean acidification is a major threat.  Examine the codes linked to corals to see what are counted as threats? -->

```{r, eval = FALSE}
cc_oa_df <- cc_sens_df %>%
  filter(stressor == 'oa' & count > 0) %>%
  left_join(spp_to_include, by = c('iucn_sid', 'cat_score'))

corals <- spp_to_include %>%
  filter(str_detect(tolower(dbf_file), 'coral'))
# table(corals$cat_score)
#   0 0.2 0.4 0.6 0.8 
# 298 175 201  27   5

coral_threats <- corals %>%
  select(iucn_sid, sciname, cat_score, threats) %>%
  distinct() %>%
  left_join(spp_threats, by = 'iucn_sid') %>%
  filter(score_num >= 3) %>%
  mutate(code_num = str_extract(code, '[0-9]+\\.') %>% as.integer()) %>%
  arrange(desc(code_num), desc(code)) %>%
  mutate(txt = paste0(code, ': ', title),
         txt = str_replace(txt, ' \\(.+\\)', ''),
         txt = fct_inorder(txt)) %>%
  select(iucn_sid, txt) %>%
  distinct() %>%
  group_by(txt) %>%
  summarize(n_spp_impacted = n()) %>%
  ungroup()
  
ggplot(coral_threats, aes(x = txt, y = n_spp_impacted)) +
  ggtheme_plot() +
  geom_col() +
  labs(x = 'Threat code', y = 'Number of spp impacted') +
  coord_flip()
  
```
<!-- Note the climate threats are limited to _11.3: Temperature extremes_ and _11.4: Storms and flooding_.  It does not appear that, for corals at least, ocean acidification is counted as _11.1: habitat shifting and alteration_ as I had originally coded it.  New strategy: just count it along with temperature extremes. -->

## Check the number of spp sensitive to different types...

```{r}
spp_sens <- cc_sens_summary %>%
  filter(iucn_sid %in% spp_maps$iucn_sid) %>%
  filter(iucn_sid %in% spp_comp$iucn_sid) %>% 
  left_join(spp_risk, by = 'iucn_sid') %>%
  filter(!is.na(cat_score) & !cat_score %in% c(0, 1)) %>%
  filter(sens)

n_spp_sens <- spp_sens %>%
  .$iucn_sid %>%
  unique() %>%
  length()

drop_check_all <- spp_marine %>%
  filter(iucn_sid %in% spp_stressed$iucn_sid) %>%
  filter(!iucn_sid %in% spp_sens$iucn_sid) %>%
  left_join(spp_narrs, by = 'iucn_sid') %>%
  left_join(spp_maps, by = 'iucn_sid') %>%
  # left_join(spp_comp %>% select(-sciname), by = 'iucn_sid') %>%
  select(iucn_sid, sciname, dimension, text) %>%
  distinct()

drop_check <- drop_check_all %>%
  filter(dimension %in% c('rationale',
                          'threats')) %>%
  ### prefilter to dramatically speed up the unnest_tokens...
  # filter(str_detect(tolower(text), 'climate|temp|acid|el ni.o')) %>%
  unnest_tokens(output = sentence, input = text, 
                token = 'sentences', drop = TRUE) %>%
  filter(str_detect(sentence, 'climate|temperature|acid|el ni.o|sea level|warming|sea ice'))
```

According to this analysis so far, there are `r n_spp_sens` distinct threatened species (with maps, comprehensively assessed, etc) affected by various climate stressors.  Note that `r n_stressed` species are listed as threatened therefore a few are not being captured via specific mentions of climate impacts.  Some may be indirect (e.g. food web interactions) or not climate related at all.  See here for a sample of some that mention climate search terms but do not seem to indicate direct impacts:

`r DT::datatable(drop_check)`

```{r}
spp_per_stressor <- cc_sens_summary %>%
  filter(sens) %>%
  count(stressor) %>%
  arrange(n) %>%
  mutate(stressor = fct_inorder(stressor))
ggplot(spp_per_stressor, aes(x = stressor, y = n)) +
  ggtheme_plot() +
  geom_col() + 
  coord_flip() +
  labs(title = 'sensitive spp by stressor',
       y = 'number of species affected by stressor')
```

```{r}
stressor_per_spp <- cc_sens_summary %>%
  filter(sens) %>%
  count(iucn_sid) %>%
  mutate(n_txt = as.character(n))

ggplot(stressor_per_spp, aes(x = n_txt)) +
  ggtheme_plot() +
  geom_bar() + 
  labs(title = 'stressor sensitivities per spp',
       x = 'number of stressors on spp',
       y = 'number of species')
```
Most species impacted by climate are impacted by two stressors (generally SST and OA).

## Check for species listed as a threat but not otherwise captured

```{r}
threats  <- read_csv(here('int/spp_threats_climate.csv')) %>%
  rename(thr_sens = sens) %>%
  mutate(thr_listed = TRUE)
# iucn_sid; cat_score; code; stressor; thr_sens; thr_listed

cc_sum <- read_csv(here('int', sprintf('spp_sens_climate_%s.csv', api_version))) %>%
  rename(cc_sens = sens)
# iucn_sid; stressor; code; gear_sens

spp_comp <- read_csv(here('_data', sprintf('iucn_comp_assessed_%s.csv', api_version)))

tmp <- threats %>%
  left_join(cc_sum, by = c('iucn_sid', 'stressor', 'code')) %>%
  group_by(iucn_sid, code) %>%
  summarize(n_sens = sum(thr_sens),
            n_stressor = sum(cc_sens, na.rm = TRUE)) %>%
  ungroup()

tmp_nomatch <- tmp %>%
  filter(n_sens >= 1 & n_stressor == 0) %>%
  group_by(iucn_sid) %>%
  summarize(codes = paste(code, collapse = ', ')) %>%
  left_join(spp_comp, by = 'iucn_sid') %>%
  filter(!is.na(taxon)) %>%
  filter(taxon != 'sturgeons')
  
table(tmp_nomatch$taxon)

DT::datatable(tmp_nomatch)
```

Note that each of these is listed as being threatened by climate pressures, but no categorization of stressor was found.  That may be legitimate - category 11 includes general impacts from habitat loss so clearly may be affected by other impacts other than SLR, SST, and OA.  Inspect each of these to better understand/inform more keywords.  (Before adding checks for enso, global warming, and sea ice there were 95 spp with a category 11 threat but no match.  Of these, about half seemed legitimate instances of climate threats).