---
title: 'Combine stressor sensitivities'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here::here('common_fxns.R'))
```


# Summary

Stressor sensitivity is a very simplified version of stressor weights (see draft1 folder for complicated version).  Here, we will simply note whether a given species is sensitive to a given stressor, using the IUCN impact weights (i.e. non-negligible impact score), ranked as 0-3 (no, low, medium, high impact) for later differentiation if necessary.  For species with NA impact scores, we can imply sensitivity based on other fields, e.g. severity.  This version of sensitivity is not related to the range of exposure, as the weight calculation was doing.

Sensitivity will denote whether a species range overlapping a stressor range constitutes an impact.

# Methods

Gather processed sensitivity dataframes and combine into a single one.
```{r}

sens_land <- read_csv(here('int', sprintf('spp_sens_land_%s.csv', api_version)),
                     col_types = cols('iucn_sid' = 'i', 'code' = 'c')) %>%
  mutate(category = 'land-based')
sens_ocean <- read_csv(here('int', sprintf('spp_sens_ocean_%s.csv', api_version)),
                     col_types = cols('iucn_sid' = 'i', 'code' = 'c')) %>%
  mutate(category = 'ocean')
sens_fis <- read_csv(here('int', sprintf('spp_sens_fishing_%s.csv', api_version)),
                     col_types = cols('iucn_sid' = 'i', 'code' = 'c')) %>%
  mutate(category = 'fishing')
sens_cc  <- read_csv(here('int', sprintf('spp_sens_climate_%s.csv', api_version)),
                     col_types = cols('iucn_sid' = 'i', 'code' = 'c')) %>%
  mutate(category = 'climate')

sens_all_df <- bind_rows(sens_land, sens_ocean, sens_fis, sens_cc)

sens_all_file <- here('_output', sprintf('spp_sensitivity_%s.csv', api_version))
write_csv(sens_all_df, sens_all_file)

```

