---
title: 'Combine stressor sensitivities'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here::here('common_fxns.R'))
```


# Summary

Examine the results of the stressor sensitivity scripts.

# Methods

## Species impacted by each stressor

Plotting the number of species impacted by each stressor.  This plot includes species that:

* have an IUCN range map
* are in a comprehensively assessed taxon (and birds are seabirds)
* are threatened or near threatened (i.e. not LC or EX)
* are sensitive to at least one stressor in our study

```{r}
sens_all_file <- here('_output', sprintf('spp_sensitivity_%s.csv', api_version))

sens_all_df <- read_csv(sens_all_file, col_types = cols('iucn_sid' = 'i'))

spp_incl <- get_incl_spp() %>%
  filter(!is.na(stressor))

plot_df <- sens_all_df %>%
  filter(sens) %>%
  select(-code, -sens) %>%
  distinct() %>%
  filter(iucn_sid %in% spp_incl$iucn_sid) %>%
  group_by(stressor, category) %>%
  summarize(n_spp = n()) %>%
  ungroup() %>%
  arrange(n_spp) %>%
  mutate(category = fct_inorder(category),
         stressor = fct_inorder(stressor))

ggplot(plot_df, aes(x = stressor, y = n_spp, fill = category)) +
  ggtheme_plot() +
  geom_col() +
  coord_flip() +
  labs(x = 'Stressor',
       y = 'Number of impacted species')
```

## Number of stressor sensitivities per species

This does not indicate spatial co-occurrence of stressors, just whether a species is sensitive to one, two, or more stressors.

```{r}

spp_incl <- get_incl_spp() %>%
  filter(!is.na(stressor))

str_count_df <- spp_incl %>%
  group_by(iucn_sid) %>%
  summarize(n_str = n_distinct(stressor),
            n_cat = n_distinct(category),
            strs = paste(unique(stressor), collapse = ', ')) %>%
  ungroup()

plot_df <- str_count_df %>%
  rename(stressors = n_str, categories = n_cat) %>%
  gather(key = type, value = ct, stressors, categories) %>%
  mutate(ct = factor(ct))

ggplot(plot_df,
       aes(x = ct, fill = type)) +
  ggtheme_plot() +
  geom_bar(position = 'dodge') +
  labs(x = '# of stressors',
       y = 'spp count')
```

## Try a pairwise table to examine # of spp with pairwise combos of stressors

```{r}
library(widyr)

spp_str_incidence <- get_incl_spp() %>%
  filter(!is.na(stressor)) %>%
  select(iucn_sid, stressor) %>%
  distinct()

result_df <- pairwise_count(spp_str_incidence,
                             item = stressor, feature = iucn_sid, 
                             sort = TRUE, diag = TRUE) %>%
  rename(str1 = item1, str2 = item2)

result_mtx <- result_df %>%
  spread(key = str2, value = n) %>%
  select(-str1) %>%
  as.matrix()

### rename rows and columns to match EEZs
rownames(result_mtx) <- colnames(result_mtx)

### set NAs to zero
result_mtx[is.na(result_mtx)] <- 0

result_mtx

mtxplot_df <- result_df %>%
  mutate(n = ifelse(str2 > str1, NA, n),
         str1 = fct_rev(str1))
ggplot(mtxplot_df, aes(x = str1, y = str2)) +
  ggtheme_plot() +
  geom_point(aes(size = n)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title = element_blank())
```
