---
title: 'Map cumulative stressor footprints'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

```


# Summary

Here we create maps of the footprint of stressor categories to help discern where "zero impact" means no stressors present, vs. no sensitive species present.

# Methods

## Get stressor maps

For each category, for each year, pull in the stressor maps for that combo and stack them up.  Make a raster showing how many stressors exist in each cell.  Cells without any stressors (in that category/year) can then be later identified as a special case of zero impact.

```{r}

str_cat_df <- get_str_cats()
str_cats <- c('all', unique(str_cat_df$category))
yrs <- 2003:2013

### filename stem for sprintf'ing the info
str_ft_stem <- here('_output/rasters/stressor_ftprints/footprint_%s_%s.tif')

for(str_cat in str_cats) {
  # str_cat <- str_cats[1]
  
  for(yr in yrs) {
    # yr <- yrs[1]
    
    str_ft_file <- sprintf(str_ft_stem, str_cat, yr)
    message('Processing ', basename(str_ft_file))
  
    if(!file.exists(str_ft_file)) {
      strs <- str_cat_df %>%
        filter(category == str_cat | str_cat == 'all') %>%
        .$stressor
      
      str_rast_stem <- file.path(dir_bd_anx, 'layers', 'stressors_10km_95vc',
                                 '%s_%s_mol10km_95vc.tif')
      str_stack <- stack(sprintf(str_rast_stem, strs, yr))
      
      ### normalize all layers to 1/NA (presence/absence of stressor)
      str_stack_norm <- str_stack / str_stack
      
      if(nlayers(str_stack) > 1) {
        str_count <- calc(str_stack_norm, fun = sum, na.rm = TRUE)
      } else {
        str_count <- str_stack_norm
      }
      values(str_count)[values(str_count) == 0] <- NA
      writeRaster(str_count, str_ft_file, overwrite = TRUE)
    }
  }
}

```

## Plot 2013 layers

```{r}
for(str_cat in str_cats) {
  # str_cat <- str_cats[1]
  f <- sprintf(str_ft_stem, str_cat, 2013)
  r <- raster(f)
  plot(r, axes = F, col = hcl.colors(20),
       main = sprintf('Stressor count (%s), 2013', str_cat))
}
```

