---
title: 'Calculate species stressor impacts'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

library(raster)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

dir_str_95cv <- file.path(dir_bd_anx, 'layers/stressors_10km_95cv')
dir_spp_maps <- file.path(dir_bd_anx, 'spp_rasts_mol_2020')
dir_spp_intsx <- file.path(dir_bd_anx, 'spp_str_rasts')
```


# Summary

For each species, pull in all associated stressor intersections, as well as species range.  For each year, summarize impacted range and % impacted range for each stressor, and for cumulative.  For each year, also calculate the number of impacts per cell and save as a map of species cumulative impact.

# Methods

## Create list of species for mapping cumulative impacts

Here we will identify all species with sensitivity to one or more stressor, and a range map.

```{r setup file names and species ids for processing}
spp_incl <- get_incl_spp() %>%
  filter(!is.na(stressor))

spp_ids <- spp_incl$iucn_sid %>%
  unique() %>% sort()

ocean_area_rast <- raster(here('_spatial/ocean_area_mol.tif'))
ocean_area_df <- data.frame(ocean_prop = values(ocean_area_rast),
                            cell_id = 1:ncell(ocean_area_rast))
```

## loop over species by ID

For each species, create a dataframe summarizing cumulative impacts:

* calculate individual stressor impacts range and pct
* calculate cumulative impact
* bind 'em together and save it as .csv to annex, `spp_impacts` folder.

### set up functions

```{r cumulative impact functions}
### set up list of all intersection files and an ID for indexing
intsx_files_all <- list.files(dir_spp_intsx, recursive = TRUE, full.names = TRUE)
intsx_id_vec    <- as.integer(str_extract(basename(intsx_files_all), '[0-9]{3,}'))
str_cats   <- read_csv(here('_raw/iucn_threat_to_stressor_lookup.csv')) %>%
  filter(!is.na(stressor)) %>%
  select(stressor, category) %>%
  mutate(stressor = str_split(stressor, ';')) %>%
  unnest(stressor) %>%
  distinct()

get_intsx_files <- function(spp_id, impact_cat) {
  ### To calc across all stressors, set impact_cat = 'all'
  ### NOTE: this uses global variables to avoid relisting all the intersection
  ### files each time...
  
  # impact_cat <- 'fishing'
  # spp_id <- spp_ids[1]
  
  ### get all the intersection files for this spp in the intsx directory
  intsx_files <- intsx_files_all[intsx_id_vec == spp_id]
  
  if(impact_cat != 'all') {
    ### subset stressors by impact category
    intsx_dirs <- basename(dirname(intsx_files))
    strs <- str_cats %>% 
      filter(category == impact_cat) %>%
      .$stressor
    intsx_files <- intsx_files[intsx_dirs %in% strs]
  }
  
  return(intsx_files)
}

calc_cum_impact <- function(spp_id, 
                            ocean_area_df,
                            dir_spp_maps,
                            impacts = 'all') {
  ### dir_spp_maps to point to location of range files;
  ### dir_spp_intsx to point to location of intersection files
  ### ocean_area_df to calc range including coastal (i.e. not 100% ocean) cells
  ### impacts allows to select subsets of impacts
  cell_area_km2 <- 105.6633 ### from ocean raster

  ### assemble file name and read the range data
  spp_range_file <- file.path(dir_spp_maps,
                              sprintf('iucn_sid_%s.csv', spp_id))
  spp_range_df <- read_csv(spp_range_file, 
                           col_types = cols('cell_id' = 'i', 'presence' = 'i')) %>%
    left_join(ocean_area_df, by = 'cell_id')
  
  spp_range_km2 <- spp_range_df %>%
    filter(presence > 0 & !is.na(presence)) %>%
    summarize(range_km2 = sum(ocean_prop) * cell_area_km2) %>%
    .$range_km2
  
  intsx_files <- get_intsx_files(spp_id, impacts)
  
  spp_intsx_df <- lapply(intsx_files,
      FUN = function(x) { # x <- intsx_files[1]
        str_name <- basename(x) %>%
          str_replace_all('spp_intsx_|_[0-9].*', '')
        
        read_csv(x, col_types = cols('cell_id' = 'i',
                                     'year' = 'i')) %>%
          mutate(stressor = str_name)
      }) %>%
    bind_rows() %>%
    filter(!is.na(cell_id)) %>%
    left_join(ocean_area_df, by = 'cell_id')
  
  spp_impact_by_stressor <- spp_intsx_df %>%
    group_by(stressor, year) %>%
    summarize(impact_km2 = sum(ocean_prop) * cell_area_km2) %>%
    ungroup()
  
  spp_cum_impact <- spp_intsx_df %>%
    select(cell_id, year, ocean_prop) %>%
    distinct() %>%
    group_by(year) %>%
    summarize(impact_km2 = sum(ocean_prop) * cell_area_km2,
              stressor = 'cumulative')
  if(nrow(spp_cum_impact) == 0) {
    spp_cum_impact <- data.frame(
      stressor = 'cumulative',
      year = NA,
      impact_km2 = 0,
      range_km2 = spp_range_km2)
  }
  
  impact_df <- bind_rows(spp_impact_by_stressor, spp_cum_impact) %>%
    mutate(range_km2 = spp_range_km2,
           pct_impact = impact_km2 / range_km2,
           iucn_sid = spp_id)
  
  return(impact_df)
}

map_cum_impact <- function(spp_id, 
                           dir_spp_maps,
                           impacts = 'all') {
  ### dir_spp_maps to point to location of range files

  ### assemble file name and read the range data
  spp_range_file <- file.path(dir_spp_maps,
                              sprintf('iucn_sid_%s.csv', spp_id))
  spp_range_df <- read_csv(spp_range_file, 
                           col_types = cols('cell_id' = 'i', 'presence' = 'i'))
  
  intsx_files <- get_intsx_files(spp_id, impacts)
  
  spp_intsx_df <- lapply(intsx_files,
      FUN = function(x) { # x <- intsx_files[1]
        str_name <- basename(x) %>%
          str_replace_all('spp_intsx_|_[0-9].*', '')
        
        read_csv(x, col_types = cols('cell_id' = 'i',
                                     'year' = 'i')) %>%
          mutate(stressor = str_name)
      }) %>%
    bind_rows() %>%
    filter(!is.na(cell_id)) 

  spp_cum_impact_map <- spp_range_df %>%
    left_join(spp_intsx_df, by = 'cell_id') %>%
      ### left_join keeps totally unimpacted cells
    group_by(cell_id, year) %>%
    summarize(n_impacts = sum(!is.na(stressor))) %>%
    ungroup() %>%
    filter(!is.na(year))
    ### Note: totally unimpacted cells will have an NA for year and
    ### zero for n_impacts; cells with impacts some years and not others
    ### will only have data for impact years...
    ### dropping these observations will make for smaller files; unimpacted
    ### ranges can be added back in later from the range map files.
  
  if(nrow(spp_cum_impact_map) == 0) {
    ### for spp with no impacts in current impact group
    spp_cum_impact_map <- data.frame(cell_id = NA, 
                                     n_impacts = 0, 
                                     year = 2003:2013)
  }
  
  return(spp_cum_impact_map)
}

```

### calculate species cumulative impacts by stressor category

``` {r calc species cumulative impacts}

reload <- FALSE
### to clear old versions:
# x <- list.files(file.path(dir_bd_anx, 'spp_impacts'), full.names = TRUE)
# unlink(x)

for(impact_cat in str_cats$category %>% unique()) {
  # impact_cat <- str_cats$category %>% unique() %>% .[1]
  # for(spp_id in spp_ids) {
  tmp <- parallel::mclapply(spp_ids, mc.cores = 12,
    FUN = function(spp_id) {
      ### spp_id <- spp_ids[2]
      ### spp_id <- y[1]
      
      spp_impact_file <- file.path(dir_bd_anx, 'spp_impacts',
                                   sprintf('spp_impacts_%s_%s.csv', spp_id, impact_cat))
      if(!file.exists(spp_impact_file) | reload) {
        message('Processing spp ', spp_id, '; ', impact_cat)
        ### Calculate total species range, accounting for partial ocean cells
        impact_df <- calc_cum_impact(spp_id, ocean_area_df,
                                     dir_spp_maps,
                                     impacts = impact_cat)
        write_csv(impact_df, spp_impact_file)
      } else {
        # message('Impact df file exists: ', basename(spp_impact_file), '... skipping!')
      }
      return(paste('success:', spp_id))
    }
  )
}

```

## Calculate maps of cumulative impacts

For each species, create a map of cumulative impacts:

* `left_join` each stressor intersection to the species range by `cell_id`
* summarize number of stressors in each cell
* write out as .csv to annex, `spp_impact_rasts`

Each of these will be performed for each stressor category.  Full cumulative can be found from simply adding 'em up.

### map species cumulative impacts by stressor category

``` {r map species cumulative impacts, eval = TRUE}

reload <- FALSE
### to clear old versions:
### x <- list.files(file.path(dir_bd_anx, 'spp_impact_rasts'), full.names = TRUE)
### unlink(x)

for(impact_cat in str_cats$category %>% unique()) {
  # impact_cat <- str_cats$category %>% unique() %>% .[1]
  # for(spp_id in spp_ids) {
  tmp <- parallel::mclapply(spp_ids, mc.cores = 4,
    FUN = function(spp_id) {
      ### spp_id <- spp_ids[2]
      ### spp_id <- y[1]
      impact_rast_file <- file.path(dir_bd_anx, 'spp_impact_rasts',
                                   sprintf('spp_impact_map_%s_%s.csv', spp_id, impact_cat))
      if(!file.exists(impact_rast_file) | reload) {
        message('Mapping spp ', spp_id, '; ', impact_cat)
        ### Calculate total species range, accounting for partial ocean cells
        impact_rast <- map_cum_impact(spp_id,
                                      dir_spp_maps,
                                      impacts = impact_cat)
        write_csv(impact_rast, impact_rast_file)
      } else {
        # message('Impact map file exists: ', basename(impact_rast_file), '... skipping!')
      }
      # return(paste('success:', spp_id))
    }
  )
}

```
