---
title: 'Map impacts per taxon'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

library(raster)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

```


# Summary

Create taxa-level maps of number of species impacted per cell, per year.  Repeat to create taxa-level maps of number of species in refugia (i.e., zero impact) per cell, per year.

# Methods

* Gather species maps by taxonomic group (IUCN comprehensively assessed groups)
* read in all impact maps for species for all years, flatten to impacted/not for each year
* add maps together for a taxa-level map for each year of number of impacted species per cell
* repeat similar process for refugia.

## Match threatened species to taxonomic groups per IUCN

Identify the species included in the impacted list (using the IUCN species IDs of impact map files, from prior scripts!).  Join this with the list of mapped species, to identify which taxonomic group the species is in (based on the downloaded IUCN shapefiles).  Taxa groups not represented in the impacted species list will be dropped.

```{r identify taxa groups}

impact_map_dir <- file.path(dir_bd_anx, 'spp_impact_rasts')

### identify spp with impact maps
spp_impacted <- list.files(impact_map_dir, recursive = TRUE) %>%
  str_extract('[0-9]+') %>%
  as.integer() %>%
  unique()

### Get inclusion list; drop species with no sensitivities to stressors
spp_sens <- get_incl_spp() %>%
  filter(!is.na(stressor))

### look for dropped taxonomic groups...
comp_file <- here('_data', sprintf('iucn_comp_assessed_%s.csv', api_version))
spp_comp <- read_csv(comp_file)

taxa_dropped <- spp_comp %>%
  filter(!taxon %in% spp_sens$taxon) %>%
  .$taxon %>% unique()

```

Taxonomic groups dropped from this analysis - i.e., group does not have any species that fits the "impacted" criteria:

`r taxa_dropped`

The one vulnerable tarpon species (i.e. not LC, EX, or DD) is not sensitive to any of the stressors and is not mapped, so the whole taxon is dropped from this analysis.

``` {r setup directories and parameters}
### directories for inputs
spp_impact_map_dir   <- file.path(dir_bd_anx, 'spp_impact_rasts')
spp_range_dir        <- file.path(dir_bd_anx, 'spp_rasts_mol_2020')

### directories for outputs
taxon_impact_map_dir <- file.path(dir_bd_anx, 'taxon_impact_rasts')
### unlink(file.path(taxon_impact_map_dir, 'taxon_impact_birds_*.*'))
taxon_refugia_map_dir <- file.path(dir_bd_anx, 'taxon_refugia_rasts')
### unlink(file.path(taxon_refugia_map_dir, 'taxon_refugia_birds_*.*'))

```

## Loop over taxonomic groups to create taxa-level impact maps

These maps ignore the number of impacts occurring on each species, and count up the number of species impacted by the aggregated stressor group (climate and non-climate).

For a given taxon:

* identify full list of impacted species
* select a subgroup (~10 spp) to process at a time
    * read in all spp in group
    * flatten all spp impact maps (for impacts > 0, set "impacted" = TRUE)
        * no need to worry about refugia cells at this stage.
    * add all maps for this group: by cell and year
* for all subgroups in a taxa, add them up
    
```{r set up functions}

get_spp_vec <- function(spp_ids, i, subgp_size) {
  gp_first <- (i - 1) * subgp_size + 1
  gp_last  <- min(i * subgp_size, length(spp_ids))
  spp_id_vec <- spp_ids[gp_first:gp_last]
}

get_spp_impact <- function(spp_id, imp_file) {
  impact_map <- read_csv(imp_file, col_types = 'iii') %>%
    filter(!is.na(year) & !is.na(cell_id)) %>%
    filter(n_impacts > 0 & !is.na(n_impacts)) %>%
    select(cell_id, year) %>%
    mutate(iucn_sid = spp_id)
  return(impact_map)
}

calc_spp_refugia <- function(spp_id, rng_file, imp_file) {
  rng_map <- read_csv(rng_file, col_types = 'ii')
  imp_map <- read_csv(imp_file, col_types = 'iii') %>%
    filter(n_impacts > 0 & !is.na(cell_id) & !is.na(year))
  
  rng_map_yrs <- rng_map %>%
    crossing(year = 2003:2013)
  
  refugia_map <- rng_map_yrs %>%
    anti_join(imp_map, by = c('cell_id', 'year')) %>%
    filter(!is.na(presence) & presence != 5) %>%
    mutate(iucn_sid = spp_id)
  
  return(refugia_map)
}

combine_maps <- function(map_list) {
  combined_map_df <- bind_rows(map_list) %>%
    group_by(cell_id, year) %>%
    summarize(n_spp = sum(n_spp, na.rm = TRUE)) %>%
    ungroup()
  
  return(combined_map_df)
}

cell_id_rast <- raster(here('_spatial/cell_id_mol.tif'))

map_to_rast <- function(map_df, cell_val = 'n_spp') {
  # map_rast <- raster::subs(cell_id_rast, as.data.frame(map_df), 
  #                          by = 'cell_id', which = val)
  map_rast <- raster(cell_id_rast)
  values(map_rast)[map_df$cell_id] <- map_df[[cell_val]]
  return(map_rast)
}

taxon_gps <- spp_sens$taxon %>% unique() %>% sort()

subgp_size <- 24 ### number of spp to process at a time
year_span <- 2003:2013
reload <- FALSE
```

```{r create taxa maps non climate}

for(taxon_gp in taxon_gps) {
  ### taxon_gp <- taxon_gps[13]
  
  taxon_map_files <- file.path(taxon_impact_map_dir, 
                              sprintf('taxon_impact_%s_%s_no_cc.tif',
                                      taxon_gp, year_span))
  
  if(any(!file.exists(taxon_map_files)) | reload) {
    spp_ids_taxa <- spp_sens %>%
      filter(taxon == taxon_gp) %>%
      .$iucn_sid %>% 
      unique() %>% sort()
    
    n_subgps <- ceiling(length(spp_ids_taxa) / subgp_size)
    message('Processing impacts for ', taxon_gp, 
            ' non-climate: ', length(spp_ids_taxa), 
            ' species broken into ', n_subgps, ' groups.')
    
    ### initialize a list to store 
    taxon_impacts_list <- vector('list', length = n_subgps)
    for(i in 1:n_subgps) {
      # i <- 1
      message('  processing group ', i)
      spp_id_vec <- get_spp_vec(spp_ids_taxa, i, subgp_size)
      
      impact_maps <- parallel::mclapply(spp_id_vec,
          FUN = function(spp_id) {
            ### spp_id <- spp_id_vec[1]
            imp_file <- file.path(spp_impact_map_dir, 
                               sprintf('spp_impact_map_%s_no_cc.csv', spp_id))
            imp_map <- get_spp_impact(spp_id, imp_file)
          }, mc.cores = subgp_size) %>%
        bind_rows()
      impacts_gp <- impact_maps %>%
        group_by(cell_id, year) %>%
        summarize(n_spp = n_distinct(iucn_sid)) %>%
        ungroup()
      
      taxon_impacts_list[[i]] <- impacts_gp
    }
    
    taxon_impacts_df <- combine_maps(taxon_impacts_list) %>%
      mutate(taxon = taxon_gp)
    
    if(nrow(taxon_impacts_df) == 0) {
      taxon_impacts_df <- data.frame(cell_id = NA,
                                     year    = NA,
                                     n_spp   = NA,
                                     taxon   = taxon_gp)
    }
    
    ### write out as rasters by year
    for(i in seq_along(year_span)) {
      taxon_map_file <- taxon_map_files[i]
      taxon_impact_yr_df <- taxon_impacts_df %>%
        filter(year == year_span[i])
      taxon_impact_yr_rast <- map_to_rast(taxon_impact_yr_df)
      writeRaster(taxon_impact_yr_rast, taxon_map_file, overwrite = TRUE)
    }

  } else {
    message('Files exist... skipping!')
  }
}
  
```


```{r create taxa maps climate only}

for(taxon_gp in taxon_gps) {
  ### taxon_gp <- taxon_gps[1]
  
  taxon_map_files <- file.path(taxon_impact_map_dir, 
                              sprintf('taxon_impact_%s_%s_cc_only.tif',
                                      taxon_gp, year_span))
  
  if(any(!file.exists(taxon_map_files)) | reload) {
    spp_ids_taxa <- spp_sens %>%
      filter(taxon == taxon_gp) %>%
      .$iucn_sid %>% 
      unique() %>% sort()
    
    n_subgps <- ceiling(length(spp_ids_taxa) / subgp_size)
    message('Processing impacts for ', taxon_gp,
            ' climate only: ', length(spp_ids_taxa), 
            ' species broken into ', n_subgps, ' groups.')
    
    ### initialize a list to store 
    taxon_impacts_list <- vector('list', length = n_subgps)
    for(i in 1:n_subgps) {
      # i <- 1
      message('  processing group ', i)
      spp_id_vec <- get_spp_vec(spp_ids_taxa, i, subgp_size)
      
      impact_maps <- parallel::mclapply(spp_id_vec,
          FUN = function(spp_id) {
            imp_file <- file.path(spp_impact_map_dir, 
                               sprintf('spp_impact_map_%s_cc_only.csv', spp_id))
            imp_map <- get_spp_impact(spp_id, imp_file)
          }, mc.cores = subgp_size) %>%
        bind_rows()
      impacts_gp <- impact_maps %>%
        group_by(cell_id, year) %>%
        summarize(n_spp = n_distinct(iucn_sid)) %>%
        ungroup()
      
      taxon_impacts_list[[i]] <- impacts_gp
    }
    
    taxon_impacts_df <- combine_maps(taxon_impacts_list) %>%
      mutate(taxon = taxon_gp)
    
    if(nrow(taxon_impacts_df) == 0) {
      taxon_impacts_df <- data.frame(cell_id = NA,
                                     year    = NA,
                                     n_spp   = NA,
                                     taxon   = taxon_gp)
    }
    
    ### write out as rasters by year
    for(i in seq_along(year_span)) {
      taxon_map_file <- taxon_map_files[i]
      taxon_impact_yr_df <- taxon_impacts_df %>%
        filter(year == year_span[i])
      taxon_impact_yr_rast <- map_to_rast(taxon_impact_yr_df)
      writeRaster(taxon_impact_yr_rast, taxon_map_file, overwrite = TRUE)
    }
    
  } else {
    message('Files exist... skipping!')
  }
}
  
```

## Loop over taxonomic groups to create taxa-level refugia maps

These maps show the number of refugia species in a cell, i.e. a species that is present but not impacted at that location.  The process will be similar to above, but the calculation subtracts the impacted range from the species range, for all years.

For a given taxon:

* identify full list of impacted species
* select a subgroup (~10 spp) to process at a time
    * read in all spp in group
    * read in all spp range maps and expand across all years
    * for each spp, use an anti_join to remove impacted cells in each year
    * then combine spp to get refugia counts for this subgroup
    * repeat over all subgroups and combine
    
```{r create taxa refugia maps non climate}

for(taxon_gp in taxon_gps) {
  ### taxon_gp <- taxon_gps[1]
  
  taxon_map_files <- file.path(taxon_refugia_map_dir, 
                               sprintf('taxon_refugia_%s_%s_no_cc.tif',
                                       taxon_gp, year_span))
  
  if(any(!file.exists(taxon_map_files)) | reload) {
    spp_ids_taxa <- spp_sens %>%
      filter(taxon == taxon_gp) %>%
      .$iucn_sid %>% 
      unique() %>% sort()
    
    n_subgps <- ceiling(length(spp_ids_taxa) / subgp_size)
    message('Processing refugia for ', taxon_gp, ' non-climate: ', length(spp_ids_taxa), 
      ' species broken into ', n_subgps, ' groups.')
    
    ### initialize a list to store 
    taxon_refugia_list <- vector('list', length = n_subgps)
    for(i in 1:n_subgps) {
      # i <- 1
      message('  processing group ', i)
      spp_id_vec <- get_spp_vec(spp_ids_taxa, i, subgp_size)

      refugia_maps <- parallel::mclapply(spp_id_vec,
          FUN = function(spp_id) {
            imp_file <- file.path(spp_impact_map_dir, 
                               sprintf('spp_impact_map_%s_no_cc.csv', spp_id))
            rng_file <- file.path(spp_range_dir,
                               sprintf('iucn_sid_%s.csv', spp_id))        
            ref_map <- calc_spp_refugia(spp_id, rng_file, imp_file)
          }, mc.cores = subgp_size) %>%
        bind_rows()
      refugia_gp <- refugia_maps %>%
        group_by(cell_id, year) %>%
        summarize(n_spp = n_distinct(iucn_sid)) %>%  
        ungroup()
      
      taxon_refugia_list[[i]] <- refugia_gp
    }
    
    taxon_refugia_df <- combine_maps(taxon_refugia_list) %>%
        mutate(taxon = taxon_gp)
    
    if(nrow(taxon_refugia_df) == 0) {
      taxon_refugia_df <- data.frame(cell_id = NA,
                                     year    = NA,
                                     n_spp   = NA,
                                     taxon   = taxon_gp)
    }

    ### write out as rasters by year
    for(i in seq_along(year_span)) {
      taxon_map_file <- taxon_map_files[i]
      taxon_ref_yr_df <- taxon_refugia_df %>%
        filter(year == year_span[i])
      taxon_ref_yr_rast <- map_to_rast(taxon_ref_yr_df)
      writeRaster(taxon_ref_yr_rast, taxon_map_file, overwrite = TRUE)
    }
    
  } else {
    message('Files exist... skipping!')
  }
}
  
```

```{r create taxa refugia maps climate only}

for(taxon_gp in taxon_gps) {
  ### taxon_gp <- taxon_gps[1]
  
  taxon_map_files <- file.path(taxon_refugia_map_dir, 
                               sprintf('taxon_refugia_%s_%s_cc_only.tif',
                                       taxon_gp, year_span))
  if(any(!file.exists(taxon_map_files)) | reload) {
    spp_ids_taxa <- spp_sens %>%
      filter(taxon == taxon_gp) %>%
      .$iucn_sid %>% 
      unique() %>% sort()
    
    n_subgps <- ceiling(length(spp_ids_taxa) / subgp_size)
    message('Processing refugia for ', taxon_gp, ' climate only: ', length(spp_ids_taxa), 
      ' species broken into ', n_subgps, ' groups.')
    
    ### initialize a list to store 
    taxon_refugia_list <- vector('list', length = n_subgps)
    for(i in 1:n_subgps) {
      # i <- 1
      message('  processing group ', i)
      spp_id_vec <- get_spp_vec(spp_ids_taxa, i, subgp_size)

      refugia_maps <- parallel::mclapply(spp_id_vec,
          FUN = function(spp_id) {
            imp_file <- file.path(spp_impact_map_dir, 
                               sprintf('spp_impact_map_%s_cc_only.csv', spp_id))
            rng_file <- file.path(spp_range_dir,
                               sprintf('iucn_sid_%s.csv', spp_id))        
            ref_map <- calc_spp_refugia(spp_id, rng_file, imp_file)
          }, mc.cores = subgp_size) %>%
        bind_rows()
      refugia_gp <- refugia_maps %>%
        group_by(cell_id, year) %>%
        summarize(n_spp = n_distinct(iucn_sid)) %>%   
        ungroup()
      
      taxon_refugia_list[[i]] <- refugia_gp
    }
    
    taxon_refugia_df <- combine_maps(taxon_refugia_list) %>%
        mutate(taxon = taxon_gp)
    
    if(nrow(taxon_refugia_df) == 0) {
      taxon_refugia_df <- data.frame(cell_id = NA,
                                     year    = NA,
                                     n_spp   = NA,
                                     taxon   = taxon_gp)
    }

   ### write out as rasters by year
    for(i in seq_along(year_span)) {
      taxon_map_file <- taxon_map_files[i]
      taxon_ref_yr_df <- taxon_refugia_df %>%
        filter(year == year_span[i])
      taxon_ref_yr_rast <- map_to_rast(taxon_ref_yr_df)
      writeRaster(taxon_ref_yr_rast, taxon_map_file, overwrite = TRUE)
    }

  } else {
    message('Files exist... skipping!')
  }
}
  
```

## Animate select taxa impact maps 

```{r animate function}

library(animation)

make_ramp <- function(n, palette) {
  
  n_even <- round(n + 50, -2)
  breaks <- seq(.5, n_even, length.out = 100)
  labels <- seq(0, n_even, by = 50)

  colors <- hcl.colors(length(breaks), palette = palette)
  
  ### For breaks and colors, add a different color for 
  ### a value of exactly zero
  return(list('colors' = c('grey40', colors),
              'breaks' = c(0, breaks),
              'labels' = labels))
}

make_gifs <- function(map_stack, filename, layer_names = NULL) {
  if(is.null(layer_names)) layer_names = names(map_stack)

  n <- max(maxValue(map_stack))
  if(is.infinite(n) | is.na(n)) n <- 1
  
  ramp <- make_ramp(n, palette = 'viridis')
  
  capture.output({
    saveGIF({
      for(i in 1:nlayers(map_stack)){
        plot(map_stack[[i]], 
             col = ramp$colors,
             breaks = ramp$breaks,
             axes = FALSE,
             axis.args = list(at = ramp$labels),
             main = layer_names[i])
      }}, 
      interval = 0.5, movie.name = filename, 
      ani.width = 700, ani.height = 420)
  })
  
  return(invisible(NULL))
}

```

```{r impact maps by taxa}

cell_id_rast <- raster(here('_spatial/cell_id_mol.tif'))
ocean_a_rast <- raster(here('_spatial/ocean_area_mol.tif'))

mapfiles <- list.files(taxon_impact_map_dir, full.names = TRUE)

all_taxa_maps <- data.frame(f = mapfiles) %>%
  mutate(type = ifelse(str_detect(f, 'no_cc.tif'), 'non-climate', 'climate only'),
         year = str_extract(basename(f), '[0-9]{4}') %>% as.integer(),
         taxon = str_replace_all(basename(f), 'taxon_impact_|_[0-9].+', ''))
taxa <- c('sea_turtles', 'mammals', 'birds', 'sharks_and_rays')

```

``` {r animate non climate, results = 'asis'}
### Animate the results
imp_type <- 'non-climate'

for(t in taxa) {
  
  gif_file <- here(sprintf('figs/impact_%s_no_cc.gif', t))
  
  if(!file.exists(gif_file) | reload) {
    # t <- taxa[1]
    taxon_map_df <- all_taxa_maps %>%
      filter(taxon == t) %>%
      filter(type == imp_type)
    rast_files <- taxon_map_df$f
    map_stack <- stack(rast_files)
    
    make_gifs(map_stack, 
              filename = gif_file,
              layer_names = paste(t, imp_type, 'impacts', taxon_map_df$year))
  }
  
  cat(sprintf('![](%s)', gif_file))
}
```

``` {r animate climate, results = 'asis'}
### Animate the results
imp_type <- 'climate only'

for(t in taxa) {
  
  gif_file <- here(sprintf('figs/impact_%s_cc_only.gif', t))
  
  if(!file.exists(gif_file) | reload) {
    # t <- taxa[1]
    taxon_map_df <- all_taxa_maps %>%
      filter(taxon == t) %>%
      filter(type == imp_type)
    rast_files <- taxon_map_df$f
    map_stack <- stack(rast_files)
    
    make_gifs(map_stack, 
              filename = gif_file,
              layer_names = paste(t, imp_type, 'impacts', taxon_map_df$year))
  }
  
  cat(sprintf('![](%s)\n', gif_file))
}
```
