---
title: 'Map richness per taxon'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

```


# Summary

To enable percentage/proportion comparisons at the taxa level, we need to know the distribution of richness of threatened/near threatened species in each taxon, including those thr/n-thr spp that are not sensitive to any of our stressors (to match methods at the full spp group level).

# Methods

* Gather species maps by taxonomic group (IUCN comprehensively assessed groups)
* read in all range maps for spp in the taxonomic group; compute a species richness and range-prioritized spp richness map for each group.

## Match threatened species to taxonomic groups per IUCN

Identify the species included in the impacted list.

```{r identify taxa groups}

### Get inclusion list; keep species with no sensitivities to stressors
spp_incl <- get_incl_spp() %>%
  left_join(get_spp_taxon(), by = c('iucn_sid', 'assess_gp')) %>%
  mutate(taxon = tolower(rank))

### directories for inputs
spp_range_dir <- file.path(dir_bd_anx, 'spp_rasts_mol_2020')

### directories for outputs
taxon_rich_map_dir <- here('_output/rasters/taxon_richness_maps')
### unlink(file.path(taxon_rich_map_dir, '*.*'))

cell_id_rast <- raster(here('_spatial/cell_id_mol.tif'))

```

## Loop over taxonomic groups to create taxa-level richness maps
    
```{r set up functions and dataframes}

get_spp_vec <- function(spp_ids, i, subgp_size) {
  gp_first <- (i - 1) * subgp_size + 1
  gp_last  <- min(i * subgp_size, length(spp_ids))
  spp_id_vec <- spp_ids[gp_first:gp_last]
}

get_spp_rangemap <- function(spp_id) {
  range_map_filestem <- file.path(spp_range_dir, 'iucn_sid_%s.csv')
  rangemap <- read_csv(sprintf(range_map_filestem, spp_id),
                       col_types = 'ii') %>%
    mutate(iucn_sid = spp_id)
  return(rangemap)
}

```

```{r create taxa maps}

### load priorities; without normalizing, priority is dependent only on spp range,
### not dependent on the set of spp considered (e.g. normalizing by max priority)
priority_df <- get_spp_priority(risk_wt = FALSE)

taxon_gps <- spp_incl$taxon %>% unique() %>% sort()

reload <- FALSE
subgp_size <- 24 ### number of spp to process at a time

for(taxon_gp in taxon_gps) {
  ### taxon_gp <- taxon_gps[1]
  
  tx_nspp_f <- file.path(taxon_rich_map_dir,
    sprintf('taxon_nspp_%s.tif', taxon_gp))
  tx_priority_sum_f <- file.path(taxon_rich_map_dir,
    sprintf('taxon_priority_sum_%s.tif', taxon_gp))

  if(any(!file.exists(tx_nspp_f, tx_priority_sum_f)) | reload) {
    spp_ids_taxa <- spp_incl %>%
      filter(taxon == taxon_gp) %>%
      .$iucn_sid %>% 
      unique() %>% sort()
    
    n_subgps <- ceiling(length(spp_ids_taxa) / subgp_size)
    message('Processing richness for ', taxon_gp, 
            ': ', length(spp_ids_taxa), 
            ' species broken into ', n_subgps, ' groups.')
    
    if(length(spp_ids_taxa) == 0) {
      ### uh oh, no species in this taxon affected by this stressor.
      taxon_rich_df <- data.frame()
    } else {
      ### loop over all spp in this taxon affected by this stressor
      ### initialize a list to store 
      taxon_rich_list <- vector('list', length = n_subgps)
      for(i in 1:n_subgps) {
        # i <- 1
        message('  processing group ', i)
        spp_id_vec <- get_spp_vec(spp_ids_taxa, i, subgp_size)
        
        ### get spp ranges and join with priority
        spp_ranges <- parallel::mclapply(spp_id_vec,
          FUN = get_spp_rangemap,
          mc.cores = subgp_size) %>%
          bind_rows() %>%
          dt_join(priority_df, by = 'iucn_sid', type = 'left') %>%
          distinct()
        
        rich_gp_summary <- spp_ranges %>%
          group_by(cell_id) %>%
          summarize(priority_sum = sum(priority, na.rm = TRUE),
                    n_spp = n()) %>%
          ungroup()
        
        taxon_rich_list[[i]] <- rich_gp_summary
      } ### end of for loop
      
    ### collate all the list element dataframes
    taxon_rich_df <- bind_rows(taxon_rich_list) %>%
      group_by(cell_id) %>%
      summarize(priority_sum = sum(priority_sum, na.rm = TRUE),
                n_spp = sum(n_spp, na.rm = TRUE)) %>%
      ungroup() %>%
      mutate(taxon = taxon_gp)
    
    } ### end of if statement for zero-length species vector
    
    if(nrow(taxon_rich_df) == 0) {
      ### means no spp with valid ranges in this list; shouldn't happen here
      stop('spp group ', taxon_gp, ' has no spp with valid range')
      taxon_rich_df <- data.frame(cell_id = NA,
                                   priority_sum = NA,
                                   n_spp = NA,
                                   taxon   = taxon_gp)
    }
    
    ### convert to rasters and write out as +/- by count and priority

    nspp_rast <- map_to_rast(taxon_rich_df, cell_val = 'n_spp')
    priority_rast <- map_to_rast(taxon_rich_df, cell_val = 'priority_sum')

    writeRaster(nspp_rast, tx_nspp_f, overwrite = TRUE)
    writeRaster(priority_rast, tx_priority_sum_f, overwrite = TRUE)

  } else {
    message('Files exist... skipping!')
  }
}
```

## Plot taxa richness maps 


```{r richness maps by taxa, eval = TRUE}

mapfiles <- list.files(taxon_rich_map_dir, full.names = TRUE)

fstem <- file.path(taxon_rich_map_dir, 'taxon_%s_%s.tif')
  ### first is type, second is taxon

for(t in taxon_gps) {
  # t <- taxon_gps[1]

  nspp_map <- raster(sprintf(fstem, 'nspp', t))
  plot(nspp_map, 
       col = hcl.colors(n = 20, rev = TRUE),
       main = sprintf('Spp richness, %s', t))
  
  priority_map <- raster(sprintf(fstem, 'priority_sum', t))
  plot(priority_map, 
       col = hcl.colors(n = 20, rev = TRUE),
       main = sprintf('Range-priority richness, %s', t))

}
```
