---
title: 'Map trends for all impacted species'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

library(raster)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

reload <- FALSE
```


# Summary

Create a species richness map for threatened marine species.

Combine taxa-level trend maps to create maps of number of threatened species overlapping intensifying or deintensifying stressors, per cell.

Plot both by absolute numbers of species impacted in a location and by proportion of species (impact counts / species richness).

This is all done at the level of cumulative impact category, not individual stressors.

# Methods

## Create species richness map for this assessment

Set up the dataframe of species to include.  By taxon, load all species range maps, smash down to taxon-level species richness; then combine these for a total species richness.

```{r Identify species in this assessment}

spp_incl <- get_incl_spp()

spp_sens <- spp_incl %>%
  filter(!is.na(stressor))

```


``` {r load id-ocean-species richness-priority maps}

cell_id_rast <- raster(here('_spatial/cell_id_mol.tif'))
ocean_a_rast <- raster(here('_spatial/ocean_area_mol.tif'))

spp_rich_rast <- raster(here('_output/rasters/n_spp_map.tif'))
values(spp_rich_rast)[values(spp_rich_rast) == 0] <- NA

# priority_sum_rast <- raster(here('_output/rasters/priority_sum_map.tif'))
# values(priority_sum_rast)[values(priority_sum_rast) == 0] <- NA

```

See the impact map script for comparisons of summed priority to summed species count...

## Setup for mapping across all taxa

For each impact category, sum increasing and decreasing stressor locations for both count and priority:

* gather all taxa-level map files for increasing or decreasing stressors into a raster stack
* collapse the stack using `calc(fun = sum, na.rm = TRUE)`

### Calculate intensification trends by count

Pull the taxon-level maps for each impact category, for both intensifying and de-intensifying impacts.

```{r trend all species by stressor category}

str_cats <- c('fishing', 'ocean', 'land-based', 'climate', 'all')

taxon_tr_dir <- file.path(dir_bd_anx, 'taxon_trend_rasts')

### Set up filename stems for this run (non-priority)
tr_map_stem <- 'taxon_trend_.+_%s_%s.tif'
  ### .+ is taxon, sprintf args are stressor category, incr/decr
outfile_stem <- here('_output/rasters/intens_maps/intens_%s_%s.tif')
  ### stressor category, incr/decr

for(str_cat in str_cats) {
  # str_cat <- str_cats[1]
  
  for(trend_direction in c('incr', 'decr')) {
    # trend_direction <- 'incr'
    
    outfile <- sprintf(outfile_stem, str_cat, trend_direction)
    message('processing ', basename(outfile))
    if(!file.exists(outfile) | reload) {
      trend_map_pattern <- sprintf(tr_map_stem, str_cat, trend_direction)
      trend_map_files <- list.files(taxon_tr_dir,
                                    pattern = trend_map_pattern,
                                    full.names = TRUE)
      trend_stack <- stack(trend_map_files)
      trend_rast <- raster::calc(trend_stack, fun = sum, na.rm = TRUE) %>%
        mask(ocean_a_rast)
      
      writeRaster(trend_rast, outfile, overwrite = TRUE)
      
    } else {
      message('File exists: ', outfile, '... skipping!')
    }
    
    trend_rast <- raster(outfile)
    trend_pct_rast <- trend_rast / spp_rich_rast
    if(any(values(trend_pct_rast) > 1, na.rm = TRUE)) {
      stop('some pct raster values > 1, fix it')
    }
    plot(trend_pct_rast,
         col = c('grey80', hcl.colors(20, palette = 'viridis')),
         axes = FALSE,
         main = sprintf('Pct species with %s %s stressors', trend_direction, str_cat))
  }
}

```

<!-- ### Calculate intensification trends by priority-weighted count -->


```{r priority-weighted trend all species by stressor category, eval = FALSE}

### Set up filename stems for this run (non-priority)
tr_map_stem <- 'taxon_priority_trend_.+_%s_%s.tif'
  ### .+ is taxon, sprintf args are stressor category, incr/decr
outfile_stem <- here('_output/rasters/intensification_priority_map_%s_%s.tif')
  ### stressor category, incr/decr

for(str_cat in str_cats) {
  # str_cat <- str_cats[1]

  for(trend_direction in c('incr', 'decr')) {
    # trend_direction <- 'incr'
    
    outfile <- sprintf(outfile_stem, trend_direction, str_cat)
    
    if(!file.exists(outfile) | reload) {
      trend_map_pattern <- sprintf(tr_map_stem, str_cat, trend_direction)
      trend_map_files <- list.files(taxon_tr_dir,
                                    pattern = trend_map_pattern,
                                    full.names = TRUE)
      trend_stack <- stack(trend_map_files)
      trend_rast <- raster::calc(trend_stack, fun = sum, na.rm = TRUE) %>%
        mask(ocean_a_rast)
      
      writeRaster(trend_rast, outfile, overwrite = TRUE)
      
    } else {
      message('File exists: ', outfile, '... skipping!')
    }
    trend_rast <- raster(outfile)
    trend_pct_rast <- trend_rast / spp_rich_rast
    plot(trend_pct_rast,
         col = c('grey80', hcl.colors(20, palette = 'viridis')),
         main = sprintf('Pct priority-wt spp with %s %s stressors', 
                        trend_direction, str_cat))
  }
}

```

### Plot net intensification

Net intensification is the difference between the number of species under intensification and the number under abatement.

```{r trend diff all species by stressor category}

str_cats <- c('fishing', 'ocean', 'land-based', 'climate', 'all')

infile_stem <- here('_output/rasters/intens_maps/intens_%s_%s.tif')
  ### stressor category, incr/decr

for(str_cat in str_cats) {
  # str_cat <- str_cats[1]

  incr <- raster(sprintf(infile_stem, str_cat, 'incr'))
  decr <- raster(sprintf(infile_stem, str_cat, 'decr'))
  
  diff <- calc(stack(incr, -decr), fun = sum, na.rm = TRUE)
  
  diff_pct <- diff / spp_rich_rast
  diff_pct[incr == 0 & decr == 0] <- NA
      
  hcl.pals('diverging')
  brks <- seq(-1, 1, .05)
  clrs <- hcl.colors(n = length(brks), palette = 'Red-Green', rev = TRUE)
  plot(diff_pct,
       col = clrs,
       breaks = brks,
       axes = FALSE,
       main = sprintf('Net intensification, Pct species, %s stressors', str_cat))

}

```

### Plot intens vs abate

Because a zero net intensification can result from low intensification cooccuring with low abatement, or high of each, let's examine the frequency with which various combinations occur.

```{r diff plot}

str_cats <- c('fishing', 'ocean', 'land-based', 'climate', 'all')

infile_stem <- here('_output/rasters/intens_maps/intens_%s_%s.tif')
  ### stressor category, incr/decr

for(str_cat in str_cats) {
  # str_cat <- str_cats[5]

  incr <- raster(sprintf(infile_stem, str_cat, 'incr'))
  decr <- raster(sprintf(infile_stem, str_cat, 'decr'))
  
  df <- data.frame(incr = values(incr), 
                   decr = values(decr), 
                   nspp = values(spp_rich_rast)) %>%
    filter(!is.na(nspp)) %>%
    filter(!is.na(incr) & incr != 0 & !is.na(decr) & decr != 0) ### we're interested in co-occurrence
  ### only 57989 cells where increase and decrease co-occur?
  
  df1 <- sample_frac(df, size = 1)

  tmp <- ggplot(df1, aes(x = decr, y = incr)) +
    geom_bin2d(bins = 15) +
    geom_abline(slope = 1, intercept = 0, color = 'red') +
    scale_fill_viridis_c() +
    scale_x_log10(breaks = c(1, 2, 5, 10, 20, 50, 100, 200)) +
    scale_y_log10(breaks = c(1, 2, 5, 10, 20, 50, 100, 200)) +
    labs(title = sprintf('Intensifying vs abating, %s stressors', str_cat),
         x = 'nspp abating', y = 'nspp intensifying')

  print(tmp)
}

```
