---
title: 'Fig SI: impacts maps by stressor and taxa'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)
library(cowplot)
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')
source(here('common_fxns.R'))
reload <- FALSE
```


# Summary

Create figure for SI: map of impacted spp by count, map of # of species.

# Methods

# SI: At-Risk Species Richness

``` {r figure SI at-risk species richness}

xfm <- function(x) log10(x) ### how are we transforming count?

nspp <- raster(here('_output/rasters/n_spp_map.tif')) 
nspp_df <- rasterToPoints(nspp) %>%
  as.data.frame() %>%
  rename(nspp = 3) %>%
  mutate(xfm_n = xfm(nspp))

max_ct <- maxValue(nspp)

ct_labels <- c(1, 10, 100, max_ct)
ct_breaks <- xfm(ct_labels)

land_sf <- read_sf(here('_spatial/ne_10m_land/ne_10m_land_no_casp.shp')) %>%
  st_transform(crs(nspp))

nspp_map <- ggplot() +
  ggtheme_map(base_size = 9) +
  geom_raster(data = nspp_df, aes(x, y, fill = xfm_n), size = .1) +
  geom_sf(data = land_sf, fill = 'grey96', color = 'grey40', size = .10) +
  scale_fill_viridis_c(breaks = ct_breaks, labels = ct_labels) +
  theme(plot.margin = unit(c(.05, 0, .05, 0), units = 'cm'),
        legend.key.width = unit(.25, 'cm')) +
  coord_sf(datum = NA) + ### ditch graticules
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = 'Species\nrichness')

# nspp_map
fname <- here('ms_figs/figS1_spp_richness.png')
ggsave(plot = nspp_map, filename = fname,
       height = 2.7, width = 6.5, units = 'in', dpi = 300)

knitr::include_graphics(fname)
```

# SI: Maps of pct impact by stressor

Load rasters of impacted spp and species count for 2013; generate impacted spp by pct; make a cool map.

``` {r figure si impact by stressor, results = 'asis'}

figfile <- here('ms_figs/figS2_impact_pct_four_stressors.png')

if(!file.exists(figfile)) {

  nspp <- raster(here('_output/rasters/n_spp_map.tif')) 
  land_sf <- read_sf(here('_spatial/ne_10m_land/ne_10m_land_no_casp.shp')) %>%
    st_transform(crs(nspp))
  stressors <- c('fishing', 'climate', 'land-based', 'ocean')
  panel_list <- vector('list', length = 4)

  for(i in seq_along(stressors)) { ### i <- 3
    
    str <- stressors[i]
    
    imp_rast_stem <- here('_output/rasters/impact_maps/impact_%s_2013.tif')
    imp_count  <- raster(sprintf(imp_rast_stem, str))
  
    imp_pct <- imp_count / nspp
  
    imp_pct_df <- rasterToPoints(imp_pct) %>%
      as.data.frame() %>%
      rename(imp_pct = layer)
  
    str_imp_map <- ggplot() +
      ggtheme_map(base_size = 9) +
      geom_raster(data = imp_pct_df, aes(x, y, fill = imp_pct)) +
      scale_fill_viridis_c(breaks = seq(0, 1, .25), 
                           labels = paste0(seq(0, 100, 25), '%'),
                           na.value = 'grey80') +
      geom_sf(data = land_sf, fill = 'grey96', color = 'grey40', size = .10) +
      theme(plot.margin = unit(c(.05, 0, .05, 0), units = 'cm'),
            legend.background = element_blank(),
            legend.key.width = unit(.25, 'cm'),
            legend.title = element_blank()) +
      coord_sf(datum = NA) + ### ditch graticules
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0))
    
    if(str == 'land-based') {
      pt_df <- imp_pct_df %>%
        filter(imp_pct > 0) %>%
        arrange(imp_pct)
      
      str_imp_map <- str_imp_map +
        geom_tile(data = pt_df, aes(x, y, fill = imp_pct, color = imp_pct), 
                  size = .2, show.legend = FALSE) +
        # geom_point(data = pt_df, aes(x = x_pt, y = y_pt, color = imp_pct),
        #            size = .05, show.legend = FALSE) +
        geom_sf(data = land_sf, fill = 'grey96', color = 'grey90', size = .05) +
        scale_color_viridis_c(breaks = seq(0, 1, .25), 
                             labels = paste0(seq(0, 100, 25), '%'))
    }
      
    panel_list[[i]] <- str_imp_map
      
  }
  
  p_a <- get_panel(panel_list[[1]])
  p_b <- get_panel(panel_list[[2]])
  p_c <- get_panel(panel_list[[3]])
  p_d <- get_panel(panel_list[[4]])
  lgd <- get_legend(panel_list[[4]])
  
  # ggdraw() + draw_plot(p_d)
  
  fig_si <- ggdraw() +
    draw_plot(p_a,  x = .01, y = 0.51, width = .42, height = .45) +
    draw_plot(p_b,  x = .45, y = 0.51, width = .42, height = .45) +
    draw_plot(p_c,  x = .01, y = 0.01, width = .42, height = .45) +
    draw_plot(p_d,  x = .45, y = 0.01, width = .42, height = .45) +
    draw_plot(lgd, x = .90, y = 0.01, width = .05, height = 1) +
    draw_label('A', x = .01, y = .99, hjust = 0, vjust = 1, 
               size = 9, color = 'grey20', fontface = 'bold') +
    draw_label('B', x = .45, y = .99, hjust = 0, vjust = 1, 
               size = 9, color = 'grey20', fontface = 'bold') +
    draw_label('C', x = .01, y = .49, hjust = 0, vjust = 1, 
               size = 9, color = 'grey20', fontface = 'bold') +
    draw_label('D', x = .45, y = .49, hjust = 0, vjust = 1, 
               size = 9, color = 'grey20', fontface = 'bold') +
    draw_label('Fishing impacts',
               x = .22, y = .99, hjust = .5, vjust = 1, 
               size = 6, color = 'grey20', fontface = 'bold') +
    draw_label('Climate impacts',
               x = .67, y = .99, hjust = .5, vjust = 1, 
               size = 6, color = 'grey20', fontface = 'bold') +
    draw_label('Land-based impacts',
               x = .22, y = .49, hjust = .5, vjust = 1, 
               size = 6, color = 'grey20', fontface = 'bold') +
    draw_label('Ocean-based impacts',
               x = .67, y = .49, hjust = .5, vjust = 1, 
               size = 6, color = 'grey20', fontface = 'bold') +
    draw_label('Species impacted\n(percent)',
               x = .93, y = .80, hjust = .5, vjust = 1, 
               size = 6, color = 'grey20', fontface = 'bold')


  ggsave(plot = fig_si, filename = figfile,
         height = 2.9, width = 6.5, units = 'in', dpi = 300)
  
}
knitr::include_graphics(figfile)

```

NOTE: Land-based pixels are exaggerated in size for display purposes

# SI: Maps of pct impact by taxon

Load rasters of impacted spp and species count for 2013; generate impacted spp by pct; make a cool map.  Let's try this with faceting instead of plotting each map separately.

``` {r figure si impact by taxon, results = 'asis'}

figfile <- here('ms_figs/figS4_impact_pct_by_taxon.png')
# unlink(figfile)

if(!file.exists(figfile)) {

  land_sf <- read_sf(here('_spatial/ne_10m_land/ne_10m_land_no_casp.shp')) %>%
    st_transform(crs(nspp))
  taxa_gps <- get_incl_spp() %>%
    filter(!is.na(stressor)) %>%
    .$desc %>% unique()
  tx_map_list <- vector('list', length = length(taxa_gps))

  for(i in seq_along(taxa_gps)) { ### i <- 1
    
    taxon <- taxa_gps[i]
    
    taxon_short <- str_replace_all(taxon, ' ', '')
    
    imp_rast_stem <- file.path(dir_bd_anx, 'taxon_impact_rasts/taxon_impact_%s_2013_all.tif')
    imp_count  <- raster(sprintf(imp_rast_stem, taxon_short))
    nspp_rast_stem <- here('_output/rasters/taxon_richness_maps/taxon_nspp_%s.tif')
    tx_nspp <- raster(sprintf(nspp_rast_stem, taxon_short))
  
    imp_pct <- imp_count / tx_nspp
  
    imp_pct_df <- rasterToPoints(imp_pct) %>%
      as.data.frame() %>%
      rename(imp_pct = layer) %>%
      mutate(desc = taxon) %>%
      arrange(imp_pct)
      
    tx_map_list[[i]] <- imp_pct_df
      
  }
  
  tx_map_df <- tx_map_list %>%
    bind_rows() %>%
    filter(!desc %in% c('cone snails', 'hagfishes'))
  
  ### find upper left coordinates for label:
  tx_label_df <- tx_map_df %>%
    select(x, y, desc) %>%
    mutate(x = min(x), y = max(y)) %>%
    distinct()
  
  ### create a dataframe of NA values to create a background
  ocean_a_rast <- raster(here('_spatial/ocean_area_mol.tif'))
  ocean_a_df <- rasterToPoints(ocean_a_rast) %>%
    as.data.frame()
  
  pt_df <- tx_map_df %>%
    filter(desc %in% c('reef-building corals', 'cone snails', 
                       'hagfishes', 'mangrove plants', 'seagrasses'))

  tx_imp_map <- ggplot(tx_map_df) +
    ggtheme_map(base_size = 9) +
    ### add grey background fill:
    geom_raster(data = ocean_a_df, aes(x, y), fill = 'grey80') +
    geom_raster(aes(x, y, fill = imp_pct)) +
    ### add geom_tile to allow for expanding pixels by adding border size
    geom_tile(data = pt_df, 
              aes(x, y, fill = imp_pct, color = imp_pct, size = coastal),
              size = .2, show.legend = FALSE) +
    scale_fill_viridis_c(breaks = seq(0, 1, .25), 
                         labels = paste0(seq(0, 100, 25), '%'),
                         na.value = 'grey80') +
    scale_color_viridis_c(na.value = 'grey80') +
    geom_sf(data = land_sf, fill = 'grey96', color = 'grey40', size = .05) +
    theme(plot.margin = unit(c(.05, 0, .05, 0), units = 'cm'),
          strip.background = element_blank(), ### drop the facet strip
          strip.text = element_text(face = 'bold'),
          strip.placement = 'inside',
          legend.background = element_blank(),
          legend.title = element_text(vjust = 1, color = 'grey20', face = 'bold'),
          legend.position = 'bottom',
          legend.key.height = unit(.15, 'cm')) +
    labs(fill = 'Percent of species impacted') +
    coord_sf(datum = NA) + ### ditch graticules
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    facet_wrap(~ desc, ncol = 2)
  
  ggsave(plot = tx_imp_map, filename = figfile,
         height = 7.6, width = 6.5, units = 'in', dpi = 300)

}
knitr::include_graphics(figfile)

```

NOTE: Land-based pixels are exaggerated in size for display purposes. Cone snails and hagfishes were dropped since the impacted areas are tiny


