---
title: 'Examine co-occurrence of impacted taxonomic groups'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)
library(widyr) ### for function: pairwise_count()
library(ggforce) ### for geom_arc_bar()
library(cowplot)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

source(here('common_fxns.R'))

reload <- FALSE
```


# Summary

Examine co-occurrence of impacts on taxonomic groups for each stressor category.  For each group, identify cells with significant impacts (e.g. greater than 25% of spp in the group are impacted), convert to a presence/absence of significant impacts, and then create a pairwise plot to show which taxonomic groups co-occur (in pairs).

Do the same for intensification maps (don't worry about abatement?)

# Methods

For each stressor category:

* pull in impact rasters for all taxonomic groups for 2013.
* flatten to "significantly impacted" by taxon.
* combine into a dataframe; join with ocean area.
* calculate pairwise impacted area for each taxon pair.

## Helper functions

* Function `get_taxon_impacts()` to read in a taxon map and convert to a data.frame with impacted spp, total spp, and pct impacted.
* Function `flatten_impacts()` to get taxon impacts (using the function above) and return only those cell IDs that meet or exceed a given threshold.

```{r}
dir_tx_imp  <- file.path(dir_bd_anx, 'taxon_impact_rasts')
dir_tx_nspp <- here('_output/rasters/taxon_richness_maps')

get_taxon_impacts <- function(tx, str_cat) {
  ### tx <- taxa_gps[1]; str_cat <- 'all'
  tx_short <- str_replace_all(tx, ' ', '')
  imp_f  <- file.path(dir_tx_imp,  sprintf('taxon_impact_%s_2013_%s.tif', tx_short, str_cat))
  nspp_f <- file.path(dir_tx_nspp, sprintf('taxon_nspp_%s.tif', tx_short))
  
  imp_r  <- raster(imp_f)
  nspp_r <- raster(nspp_f)
  imp_df <- data.frame(cell_id = 1:ncell(imp_r),
                       imp  = values(imp_r),
                       nspp = values(nspp_r)) %>%
    filter(!is.na(imp) & !is.na(nspp)) %>%
    mutate(desc = tx)
}

flatten_impacts <- function(tx, str_cat, thresh) {
  sig_imp_df <- get_taxon_impacts(tx, str_cat) %>%
    filter((imp / nspp) >= thresh) %>%
    select(cell_id, desc) %>%
    mutate(stressor = str_cat)
}
```

```{r pairwise plot for stressor categories}

ocean_a_rast <- raster(here('_spatial/ocean_area_mol.tif'))
ocean_a_df <- data.frame(cell_id = 1:ncell(ocean_a_rast),
                         a_prop = values(ocean_a_rast)) %>%
  filter(!is.na(a_prop)) %>%
  mutate(a_km2 = a_prop * res(ocean_a_rast)[1]^2 / 1e6)

spp_sens <- get_incl_spp() %>%
  filter(!is.na(stressor))

taxa_gps <- spp_sens$desc %>% unique() %>% sort()
str_cats <- c('all', 'climate', 'ocean', 'fishing', 'land-based')

### first assemble df for full-ocean impacts using 25% threshold
pairwise_results_file <- here('_output/taxa_pairwise_results.csv')

if(!file.exists(pairwise_results_file)) {
  results_df <- data.frame()
  for(str_cat in str_cats) {
    ### str_cat <- 'climate'
    
    tx_str_df <- parallel::mclapply( # tx <- taxa_gps[1]
      taxa_gps, FUN = function(tx) {
        df <- flatten_impacts(tx, str_cat, thresh = .25)
      }, mc.cores = 10) %>%
      bind_rows() %>%
      dt_join(ocean_a_df, by = 'cell_id', type = 'left')
    
    tmp_df <- pairwise_count(tx_str_df,
                                item = desc, feature = cell_id, wt = a_km2,
                                sort = TRUE, diag = TRUE) %>%
      rename(tx1 = item1, tx2 = item2, area_km2 = n) %>%
      mutate(str = str_cat)
    
    results_df <- bind_rows(results_df, tmp_df)
  }
  
  write_csv(results_df, pairwise_results_file)
}

```

``` {r}

results_df <- read_csv(pairwise_results_file)

result_order <- results_df %>%
  filter(str == 'all') %>%
  filter(tx1 == tx2) %>%
  arrange(area_km2) %>%
  .$tx1

### function to transform area into a radius - for consistency
r_xfm <- function(area_km2) {
  # r <- (area_km2 / max(area_km2, na.rm = TRUE))^(1/2) / 2 + .08
  r <- log(area_km2 + 100) / log(max(area_km2, na.rm = TRUE)) / 1.8 - .05
}

### Note: geom_arc_bar doesn't seem to like factors on axes
mtxplot_df <- results_df %>%
  mutate(tx1 = factor(tx1, levels = result_order),
         tx2 = factor(tx2, levels = rev(result_order))) %>%
  mutate(tx1_n = as.integer(tx1),
         tx2_n = as.integer(tx2),
         str = factor(str),
         r = r_xfm(area_km2),
         start = (as.integer(str) - 1) * pi/2,
         end   = (as.integer(str)) * pi/2) %>%
  filter(tx1_n + tx2_n <= 11)


fcol_mtxplot_df <- mtxplot_df %>%
  filter(str != 'all')
all_mtxplot_df <- mtxplot_df %>%
  filter(str == 'all')

bkgd_tri <- data.frame(group = c(1, 1, 1), 
                       x = c(.5, 10.5, 10.5), 
                       y = c(10.5, 10.5, .5))

fig_pairwise_plot <- ggplot() +
  ggtheme_plot(base_size = 11) +
  geom_polygon(data = bkgd_tri, aes(x = x, y = y, group = group),
               fill = 'white', color = 'white') +
  geom_abline(intercept = 11, slope = -1, color = 'darkred', alpha = .5) +
  geom_circle(data = all_mtxplot_df, aes(x0 = tx1_n, y0 = tx2_n, r = r),
              size = .1, color = 'grey30', fill = 'white') +
  geom_arc_bar(data = fcol_mtxplot_df, 
               aes(x0 = tx1_n, y0 = tx2_n, r = r, 
                   r0 = 0, start = start, end = end, fill = str),
               color = 'grey30', size = .1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title = element_blank()) +
  scale_x_continuous(breaks = 1:length(taxa_gps),
                     labels = levels(mtxplot_df$tx1),
                     expand = c(0, 0)) +
  scale_y_continuous(breaks = 1:length(taxa_gps),
                     labels = levels(mtxplot_df$tx2),
                     expand = c(0, 0)) +
  coord_fixed() +
  scale_fill_viridis_d()

a_max <- results_df$area_km2 %>% max()
legend_df <- data.frame(
    x = 0, y = 0, 
    area_km2 = c(10^(c(2, 4, 6, 8)), a_max),
    str = c('land-based', 'ocean', 'climate', 'fishing', 'all')
  ) %>%
  mutate(str = factor(str, levels = levels(mtxplot_df$str)),
         r = r_xfm(area_km2),
         start = (as.integer(str) - 1) * pi/2,
         end   = (as.integer(str)) * pi/2)
  

fig_legend1 <- ggplot() +
  theme_void() +
  geom_circle(data = legend_df %>% filter(str == 'all'), 
              aes(x0 = x, y0 = y, r = r),
              size = .1, color = 'grey30', fill = 'white') +
  geom_arc_bar(data = legend_df %>% filter(str != 'all'), 
               aes(x0 = x, y0 = y, r = r, 
                   r0 = 0, start = start, end = end, fill = str),
               color = 'grey30', size = .1, show.legend = FALSE) +
  coord_fixed() +
  scale_fill_viridis_d()

fig_legend2 <- ggplot() +
  theme_void() +
  geom_circle(data = legend_df %>% filter(str == 'all'), 
              aes(x0 = x, y0 = y, r = r),
              size = .1, color = 'grey30', fill = NA) +
  geom_arc_bar(data = legend_df %>% filter(str != 'all'), 
               aes(x0 = x, y0 = y, r = r, 
                   r0 = 0, start = start, end = end),
               color = 'grey30', fill = 'grey80', size = .1, show.legend = FALSE) +
  coord_fixed() +
  scale_fill_viridis_d()

# fig_pairwise_plot
# fig_legend
```

## Combine Fig 3a pieces

```{r}

lgd1_ctr <- c(.75, .85)
lgd2_ctr <- c(.75, .72)
lgd_size <- .085 ### This matches the mammal largest circle pretty well

fig <- ggdraw() +
  ### main plot
  draw_plot(fig_pairwise_plot + theme(legend.position = 'none')) +
  ### line from legend outline to "cumulative" label
  draw_line(x = c(lgd1_ctr[1], lgd1_ctr[1] - .1 * lgd_size),
            y = c(lgd1_ctr[2], lgd1_ctr[2] + .65 * lgd_size),
            size = .1, color = 'grey30') +
  ### draw legend 1 figure
  draw_plot(fig_legend1, 
            x = lgd1_ctr[1] - .5 * lgd_size, 
            y = lgd1_ctr[2] - .5 * lgd_size, 
            width = lgd_size, height = lgd_size) +
  ### draw labels for legend 1
  draw_label('climate',  color = 'grey20',
             x = lgd1_ctr[1] + .3*lgd_size, y = lgd1_ctr[2] - .25*lgd_size,
             hjust = 0, size = 7) +
  draw_label('fishing',  color = 'grey20',
             x = lgd1_ctr[1] - .4*lgd_size, y = lgd1_ctr[2] - .25*lgd_size,
             hjust = 1, size = 7) +
  draw_label('land-based',  color = 'grey20',
             x = lgd1_ctr[1] - .2*lgd_size, y = lgd1_ctr[2] + .25*lgd_size,
             hjust = 1, size = 7) +
  draw_label('ocean',  color = 'grey20',
             x = lgd1_ctr[1] + .2*lgd_size, y = lgd1_ctr[2] + .25*lgd_size,
             hjust = 0, size = 7) +
  draw_label('cumulative',  color = 'grey20',
             x = lgd1_ctr[1], y = lgd1_ctr[2] + .67 * lgd_size,
             hjust = .5, vjust = 0, size = 7) +
  ### line from legend2 center to quadrants for area label
  draw_line(x = c(lgd2_ctr[1], lgd2_ctr[1] - .5*lgd_size),
            y = c(lgd2_ctr[2], lgd2_ctr[2] + .3*lgd_size),
            size = .2, color = 'grey30') +
  draw_line(x = c(lgd2_ctr[1], lgd2_ctr[1] + .5*lgd_size),
            y = c(lgd2_ctr[2], lgd2_ctr[2] + .3*lgd_size),
            size = .2, color = 'grey30') +
  draw_line(x = c(lgd2_ctr[1], lgd2_ctr[1] + .5*lgd_size),
            y = c(lgd2_ctr[2], lgd2_ctr[2] - .3*lgd_size),
            size = .2, color = 'grey30') +
  draw_line(x = c(lgd2_ctr[1], lgd2_ctr[1] - .5*lgd_size),
            y = c(lgd2_ctr[2], lgd2_ctr[2] - .3*lgd_size),
            size = .2, color = 'grey30') +
  draw_label('100,000,000 km²', 
             x = lgd2_ctr[1] - .55*lgd_size, y = lgd2_ctr[2] - .3*lgd_size,
            hjust = 1, size = 7, color = 'grey20') +
  draw_label('10,000 km²', 
             x = lgd2_ctr[1] + .55*lgd_size, y = lgd2_ctr[2] + .3*lgd_size,
            hjust = 0, size = 7, color = 'grey20') +
  draw_label('100 km²',
             x = lgd2_ctr[1] - .55*lgd_size, y = lgd2_ctr[2] + .3*lgd_size,
            hjust = 1, size = 7, color = 'grey20') +
  draw_label('1,000,000 km²', 
             x = lgd2_ctr[1] + .55*lgd_size, y = lgd2_ctr[2] - .3*lgd_size,
            hjust = 0, size = 7, color = 'grey20') +
  ### draw legend 2 figure
  draw_plot(fig_legend2, 
            x = lgd2_ctr[1] - .5 * lgd_size, 
            y = lgd2_ctr[2] - .5 * lgd_size, 
            width = lgd_size, height = lgd_size)

 
fig_file <- here('ms_figs/figS6_taxa_cooccurrence.png')

ggsave(plot = fig, filename = fig_file,
       width = 6.5, height = 6.5, units = 'in', dpi = 300)

knitr::include_graphics(fig_file)
```


